---
title: "Simulated Probabilistic Distributions"
author: "Win Cowger"
date: "06/3/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---


#Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r library}
# Bimodal universal shape distribution for environmental microplastic from Kooi and Koelmans (2019)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
library(drc) #dose-response curves
library(readr)
library(readxl)
library(MASS)
library(truncnorm)
library(tidyverse)
library(MonteCarlo)# Monte Carlo Simulation 
library(gridExtra)
library(msm) ## rtnorm - get upper and lower limit of shape distribution
library(GeneralizedHyperbolic) ## normal-inverse Gaussian
library(skimr)
library(drc) #dose-response curves
library(readr)
library(readxl)
library(DALEX)
library(tigerstats)
library(caret)
library(knitr)
library(modelplotr)
library(rsample)
library(interpret) #explainable boosting machine
#library(ssdtools)
library(ggrepel)
library(scales)
library(ggdark)
library(ggsci)
```

```{r Theme, include=FALSE}
#Theme type
     theme.type<- theme_bw(base_size = 14) +
                  #     dark_theme_bw(base_size = 15) +
                    theme(plot.title = element_text(hjust = 0.5),
                    plot.subtitle = element_text(hjust = 0.5))

     #color selection
     fill.type <-    #scale_fill_viridis(discrete = TRUE)#,
                         #scale_fill_brewer(palette = "Paired"),
                         # scale_fill_tron()#,
                         # scale_fill_locuszoom(),
                         # scale_fill_d3(),
                          scale_fill_npg()#,
                         # scale_fill_jama())
     #color selection
     color.type <- #scale_color_viridis(discrete = TRUE)#,
                         # scale_color_brewer(palette = "Paired"),
                          #scale_color_tron()#,
                         # scale_color_locuszoom(),
                         # scale_color_d3(),
                          scale_color_npg()#,
                         # scale_color_jama())

```

#The idea here is we want to be able to just change the dataset and model here and then the entire code base will react to the update. The challenge is that making the right synthetic and uniform datasets will differ depending on the column types. Is it safe to assume that there will be no numeric variables in the model besides the concentrations which will be present in the development of the synthetic and uniform datasets?  
```{r data import}
require(readr)
#load aoc_z into dataframe. This file is generated from RDA_Maker.R
#source("Tox Data/RDA_Maker.R")
aoc_z <- readRDS(file = "Tox Data/aoc_z.Rda")
## Data filtering
```

To ensure quality data feeds the model, we are filtering for technical red criteria.
```{r}
## First filter data with global filters
aoc_intermediate <- aoc_z %>% 
  filter(size.length.um.used.for.conversions > 1) %>%  #alignments only valid above 1 um
  filter(tier_zero_tech_f == "Red Criteria Passed",
         #tier_zero_risk_f == "Red Criteria Passed",
         polymer != "Not Reported",
         !environment %in% c("Terrestrial", "Not Reported"),
         #org_f != "Bacterium",
         org_f != "Plant",
         #effect.metric != "HONEC",
         risk.13 != 0 #Drop studies that received a score of 0 for endpoints criteria (this also drops studies that have not yet been scored) - KEEP THIS AFTER THE RED CRITERIA FILTERS  
         ) %>% 
  #Remove 26C temperature treatment data from Jaimukar et al. 2018
  filter(!(article == 42 & media.temp == 26)) %>% 
  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) %>%   #makes it less confusing below
   #tell model if data if size is translocatable or not
  #mutate(translocatable = case_when(size.length.um.used.for.conversions >1 & size.length.um.used.for.conversions< 83 ~ "translocatable",
                        #size.length.um.used.for.conversions >83 ~ "non-translocatable",
                        #size.length.um.used.for.conversions < 1 ~ "nano"
                        #            )) %>% 
  droplevels()  #eliminate polymer aand shape data that's not needed
```


The idea here is we want to be able to just change the dataset and model here and then the entire code base will react to the update. The challenge is that making the right synthetic and uniform datasets will differ depending on the column types. Is it safe to assume that there will be no numeric variables in the model besides the concentrations which will be present in the development of the synthetic and uniform datasets?  

# Shape and Polymer Composition in Database vs. Environment
## Polmyer
```{r}
require(reshape2)
Environment_poly <- tibble(
        EVA = 0.06,
        LTX = 0,
        PA = 0.12,
        PC = 0,
        PE = 0.25,
        PET = 0.165,
        PLA = 0,
        PMMA = 0,
        PP = 0.145,
        PS = 0.085,
        PUR = 0,
        PVC = 0.02
) %>% 
  melt() %>% 
  rename("polymer" = "variable",
         "freq" = "value") %>% 
  mutate(type = "environment")

# plot
aoc_intermediate %>% 
  filter(polymer != "Not Reported") %>% 
  group_by(polymer) %>% 
  summarize(count = n()) %>% 
  mutate(freq = count / sum(count)) %>% 
  mutate(polymer = fct_reorder(polymer, freq),
         type = "tox") %>% 
  dplyr::select(-count) %>% 
  rbind(Environment_poly) %>% 
  ggplot(aes(x = freq, y = polymer, fill = type)) +
  geom_col(position = "dodge") +
  labs(title = "Microplastics Polymers in Environment vs. Toxicity Database") +
  scale_x_continuous(labels = scales::percent) +
  xlab("Relative Frequency") +
  theme.type +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.position = "top") +
  scale_fill_d3(labels = c("Environment", "Toxicity Database"))
```

## Shape
```{r}
Environment_shape <- tibble(
        fiber = 0.485,
        fragment = 0.31,
        sphere = 0.065,
        film = 0.055,
        foam = 0.035
) %>% 
  melt() %>% 
  rename("shape" = "variable",
         "freq" = "value") %>% 
  mutate(type = "environment")

# plot
aoc_intermediate %>% 
  filter(shape != "Not Reported") %>% 
  group_by(shape) %>% 
  summarize(count = n()) %>% 
  mutate(freq = count / sum(count)) %>% 
  mutate(shape = fct_reorder(shape, freq),
         type = "tox") %>% 
  dplyr::select(-count) %>% 
  rbind(Environment_shape) %>% 
  ggplot(aes(x = freq, y = shape, fill = type)) +
  geom_col(position = "dodge") +
  labs(title = "Microplastics Shape Abundances in Environment vs. Toxicity Database") +
  xlab("Relative Frequency") +
  scale_x_continuous(labels = scales::percent) +
  theme.type +
  theme(plot.title = element_text(hjust = 0.5),
        legend.title = element_blank(),
        legend.position = "top") +
  scale_fill_d3(labels = c("Environment", "Toxicity Database"))

```


# Model Build

### Alignment

#### Parameters
```{r}
## parametrization ##
# Define params for correction #
alpha = 2.07 #table s4 for marine surface water. length
# define parameters for power law coefficients
a.sa = 1.5 #marine surface area power law
a.v = 1.48 #a_V for marine surface water volume
a.m = 1.32 # upper limit fora_m for mass for marine surface water in table S4 
a.ssa = 1.98 # A_SSA for marine surface water

#define additional parameters for calculations based on averages in the environment
R.ave = 0.77 #average width to length ratio for microplastics in marine enviornment
p.ave = 1.10 #average density in marine surface water

#join alpha values for each data point
aoc_intermediate_alphas <- aoc_intermediate %>% 
  mutate(alpha = alpha) %>% 
  mutate(a.sa = a.sa) %>% 
   mutate(a.v =  a.v) %>% 
   mutate(a.m =  a.m) %>% 
   mutate(a.ssa = a.ssa) %>% 
   mutate(R.ave = R.ave) %>% 
   mutate(p.ave = p.ave)
```

#### Functions
```{r}
###function to derive correction factor (CF) from Koelmans et al (equation 2)
CFfnx = function(a, #default alpha from Koelmans et al (2020)
                 x2D, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
                 x1D, #1 um is lower default size
                 x2M, x1M){
  CF = (x2D^(1-a)-x1D^(1-a))/(x2M^(1-a)-x1M^(1-a)) 
  return(CF)}

#### equations for mu_x_poly (note that there are three depending on certain alphas for limits of equation)
#generalizable if a.x =2 or not
mux.polyfnx_generalizable = Vectorize(function(a.x, x_UL, x_LL){
  if(a.x == 1){ # in case a.x = 1
    mux.poly = (x_UL - x_LL)/(log(x_UL/x_LL))
    return(mux.poly)}
  if(a.x == 2){ # in case a.x = 2
     mux.poly = (log10(x_UL/x_LL))/(x_LL^(-1) - x_UL^-1)
     return(mux.poly)}
  else{ #in case alpha is not 2 or 1
    mux.poly = ((1-a.x)/(2-a.x)) * ((x_UL^(2-a.x) - x_LL^(2-a.x))/(x_UL^(1-a.x) - x_LL^(1-a.x)))
    return(mux.poly)}
  },
  vectorize.args = "a.x") # if Vectorize isn't here, the if else won't work
## ^^ Note that the above generalizable function doesn't play well with mutate(case_when), likely due to some bug with dplyr. I don't have a solution to this, so a special equation will need to be used when those values are used...

#in case alpha is not 1 or 2
mux.polyfnx = function(a.x, x_UL, x_LL){
    mux.poly = ((1-a.x)/(2-a.x)) * ((x_UL^(2-a.x) - x_LL^(2-a.x))/(x_UL^(1-a.x) - x_LL^(1-a.x)))
    return(mux.poly)}

##### If alpha does equal 2 #####
mux.polyfnx2 = function(a.x, x_UL,x_LL){
  mux.poly = (log(x_UL/x_LL))/(x_LL^(-1) - x_UL^-1)
  return(mux.poly)}

##### If alpha equals 1 #####
mux.polyfnx1 = function(a.x, x_UL, x_LL){
     mux.poly = (x_UL - x_LL)/(log(x_UL/x_LL)) #natural log
    return(mux.poly)}

### Calculating max ingestible parameters ###
## function to calcualte min and max ingestible surface area ##
SAfnx = function(a, # a = 0.5 * length
                 b, # b = 0.5 * width
                 c # c = 0.5 * height (note that hieght is 0.67 * width)
){
  SA = 4*pi*(((a*b)^1.6 + (a*c)^1.6 + (b*c)^1.6) / 3)^(1/1.6)
  return(SA)}

## max ingestible volume ##

volumefnx = function(R, L){
  volume = 0.111667 * pi * R^2 * L^3 #assumes height = 0.67 * Width, and Width:Length ratio is 'R' (compartment-specific)
  return(volume)}

volumefnx_poly = function(width, length){
  height = width #0.67 * width
  volume = (4/3) * pi * (length/2) * (width/2) * (height/2) #assumes height = 0.67 * Width 
  return(volume)}

#max ingestible mass (only used for mu_mono calculations)
massfnx = function(R, L, p){
  mass = p * #density (g/cm^3)
    0.111667 * pi * R^2 * L^3 * # volume (um^3): assumes height = 0.67 * Width, and Width:Length ratio is 'R' (compartment-specific)
    1/1e12 * 1e6 #correction factor
  return(mass)}

massfnx_poly = function(width, length, p){
  height = width #0.67 * width
  volume = (4/3) * pi * (length/2) * (width/2) * (height/2) #assumes height = 0.67 * Width 
  mass = p * #density (g/cm^3)
    volume * # volume (um^3): assumes height = 0.67 * Width, and Width:Length ratio is 'R' (compartment-specific)
    1/1e12 * 1e6 #correction factor
  return(mass)}

#max ingestible specific surface area
SSAfnx = function(sa, #surface area, calcaulted elsewhere
                  m){ #mass, calculated elsewhere
  SSA = sa/m
    return(SSA)}

#max ingestible specific surface area
SSA.inversefnx = function(sa, #surface area, calcaulted elsewhere
                  m){ #mass, calculated elsewhere
  SSA.inverse = m / sa
    return(SSA.inverse)}
```

#### Calculate

Here we will calculate two aligned exposure concentrations: surface area (1 - 83 um), and volume (1 - 5,000 um). For both, the upper aligned value is the smaller of either the nominal size listed or the mouth size of the species.
##### Tissue Translocation 
```{r}
###define sizes for alignment##
x1M_set <- 1 #um lower size for all alignments
x1D_set <- 1 #um lower size for all alignments
x2D_set <- 5000 #um
upper.tissue.trans.size.um <- 83 #10 #um #set size for x2M

# calculate ERM for each species
aoc_final <- aoc_intermediate_alphas  %>% 
   #### TISSUE TRANSLOCATION ####
# define upper size length for Translocation 
#set to 83um for upper limit or max size ingest, whichever is smaller
mutate(x2M_trans = case_when(is.na(max.size.ingest.um) ~ upper.tissue.trans.size.um, 
                             max.size.ingest.um  < upper.tissue.trans.size.um ~  max.size.ingest.um,
                             max.size.ingest.um  > upper.tissue.trans.size.um ~ upper.tissue.trans.size.um)) %>% 
  
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL_trans = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly_trans = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M_trans, #upper ingestible size limit (width of particle)
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL_trans = (EC_mono_p.particles.mL_trans * mu.p.mono)/mu.p.poly_trans) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio_trans = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M_trans, #upper translocatable
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL_trans = EC_poly_p.particles.mL_trans * CF_bio_trans) %>%  #aligned particle effect concentraiton (1-5000 um)
  
  #### Surface area ERM ####
##--- environmental calculations ---###
  #calculate lower translocatable surface area
  mutate(x_LL_sa_trans = SAfnx(a = 0.5 * x1D_set, #length
                               b = 0.5 * x1D_set, #0.5 * R.ave * x1D_set, #width
                               c = 0.5 * x1D_set  #0.5 * R.ave * 0.67 * x1D_set #height
                               )) %>%  
  #calculate upper translocatable surface area
  mutate(x_UL_sa_trans = SAfnx(a = 0.5 * x2M_trans, 
                               b = 0.5 * x2M_trans, #width #0.5 * R.ave * x2M, 
                               c = 0.5 * x2M_trans #heigth #0.5 * R.ave * 0.67 * x2M
                               )) %>%  
  #calculate mu_x_poly (env) for surface area
  mutate(mu.sa.poly_trans = mux.polyfnx(a.sa, x_UL_sa_trans, x_LL_sa_trans)) %>% 
  
  ##--- laboratory calculations ---###
  ## define mu_x_mono OR mu_x_poly (lab) for alignment to ERM  #
  #(note that if mixed particles were used, a different equation must be used)
  mutate(mu.sa.mono = case_when(
    polydispersity == "monodisperse" ~ particle.surface.area.um2, # use reported surface area in monodisperse
    polydispersity == "polydisperse" ~  mux.polyfnx(a.x = a.sa, 
                                  x_LL = particle.surface.area.um2.min,
                                  x_UL = particle.surface.area.um2.max))) %>% 
  
   #calculate polydisperse effect concentration for surface area (particles/mL)
  mutate(EC_poly_sa.particles.mL_trans = (EC_mono_p.particles.mL_trans * mu.sa.mono)/mu.sa.poly_trans) %>%  
  #calculate environmentally realistic effect threshold
  mutate(EC_env_sa.particles.mL_trans = EC_poly_sa.particles.mL_trans * CF_bio_trans) %>% 
  
  ##### FOOD DILUTION ####
  # define upper size length for ingestion 
  mutate(x2M_ingest = case_when(is.na(max.size.ingest.um) ~ x2D_set, 
                         max.size.ingest.um < x2D_set ~ max.size.ingest.um,
                         max.size.ingest.um > x2D_set ~ x2D_set
                         )) %>%  #set to 5,000 as upper limit or max size ingest, whichever is smaller
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL_ingest = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly_ingest = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M_ingest, #upper ingestible size limit
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL_ingest = (EC_mono_p.particles.mL_ingest * mu.p.mono)/mu.p.poly_ingest) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio_ingest = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M_ingest, #upper ingestible length
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default upper size range
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL_ingest = EC_poly_p.particles.mL_ingest * CF_bio_ingest) %>%  #aligned particle effect concentraiton (1-5000 um)
  
  
  #### volume ERM ####
##--- environmental calculations ---###
  #calculate lower ingestible volume 
  mutate(x_LL_v_ingest = volumefnx_poly(length = x1D_set,
                                 width = x1D_set)) %>% 
  #calculate maximum ingestible volume 
  mutate(x_UL_v_ingest = volumefnx_poly(length = x2M_ingest, # length-limited
                                 #x2D_set, #upper definiton (accouunts for fibers) CONSERVATIVE
                                 width = x2M_ingest)) %>% #ingestion-limited
  # calculate mu.v.poly
  mutate(mu.v.poly_ingest = mux.polyfnx(a.v, x_UL_v_ingest, x_LL_v_ingest)) %>% 
  ##--- laboratory calculations ---###
  ## define mu_x_mono OR mu_x_poly (lab) for alignment to ERM  #
  #(note that if mixed particles were used, a different equation must be used)
  mutate(mu.v.mono = case_when(
    polydispersity == "monodisperse" ~ particle.volume.um3, # use reported volume in monodisperse
    polydispersity == "polydisperse" ~ mux.polyfnx(a.x = a.v, 
                                                   x_LL = particle.volume.um3.min,
                                                   x_UL = particle.volume.um3.max))) %>% 
  
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_v.particles.mL_ingest = (EC_mono_p.particles.mL_ingest * mu.v.mono)/mu.v.poly_ingest) %>%  
    #calculate environmentally realistic effect threshold
  mutate(EC_env_v.particles.mL_ingest = EC_poly_v.particles.mL_ingest * CF_bio_ingest) %>% 
  
   ###### CLEANUP #####
  mutate(particles.mL.ox.stress = EC_env_sa.particles.mL_trans,
         particles.mL.food.dilution = EC_env_v.particles.mL_ingest)
```



## Determine how important features are for predicting concentration
```{r}
test_mutual_information <- aoc_final %>%
  #filter data select predictor variables#
  dplyr::select(c(
    ### DOSE METRICS ####
                  particles.mL.ox.stress,
                  particles.mL.food.dilution,
                  dose.mg.L.master, 
                  dose.surface.area.um2.mL.master, #fibers are assumed to be cylinders, fragments are assumed to be spheres
                  dose.particles.mL.master, 
    ### Organism characteristics ###
                  organism.group, 
                  life.stage, #use adults for prediction
                  species_f, 
                  max.size.ingest.mm,
    ### Exposure Conditions ####              
                  bio.org, #use tissue and above for lower tiers, organism and above for higher tiers
                  exposure.route, #train on all, but only use "aquatic" for build
                  environment, #train on all, buit only use "marine" for SSD Build
                  acute.chronic_f, #use 'chronic' for SSD build
                  exposure.duration.d, 
    ### Particle characteristics ####
                  polymer, 
                  shape,
                  density.g.cm3,
                  size.length.um.used.for.conversions, #we can't use size because we're using mixtures of particles! Have to specify translocatable or not
                  translocatable,
    ### Effect data ####
                  effect, 
                  effect.score, #relative numerical strength of effect
                  effect.metric, #train on all, but only specify "NOEC" for SSD build
                  lvl1_f, #keep it generic for easy SSD build (use fitness)
                  lvl2_f,
                  #lvl3_f
                  )) #%>%
  drop_na()


```


```{r}

 x2y(x = test_mutual_information$dose.mg.L.master, y = test_mutual_information$organism.group)

```


```{r}
# Let's generate the data again
set.seed(42)

library(rpart)
preds <- predict(rpart(organism.group~dose.mg.L.master, data = test_mutual_information, method = "class"), type = "vector")

# Set up a chart
ggplot(data = test_mutual_information) +
  geom_point(aes(x = dose.mg.L.master, y = organism.group), size = 0.5) +
  #geom_line(aes(y=preds, color = '2')) +
  scale_x_log10()
```













# Resources: 
https://pubs.acs.org/doi/suppl/10.1021/acs.est.0c02982/suppl_file/es0c02982_si_001.pdf https://pubs.acs.org/doi/suppl/10.1021/acs.estlett.9b00379/suppl_file/ez9b00379_si_001.pdf
https://pubs.acs.org/doi/10.1021/acs.estlett.9b00379


# Particles constant, volume different
## Generate Particles

Shape is defined as:
Simplifed equation for corey shape factor:

$CSF = H/sqrt(LW)$

Equation for abundance of particles based on density (two distributions). (Note that this is *equation 4* in Kooi and Koelmans (2019).)

The most abundant shape category of microplastic in water and sediment is fibers (48.5%), followed by fragments (31%), beads (6.5%), films (5.5%), and foam (3.5%)(3). Combining these abundance data with the triangular shape distributions (Table S2 and Figure S2) resulted in a continuous bimodal microplastic shape distribution (Figure 2A). The fitted parameter values for eq 4 were as follows: f1 = 0.06, f2 = 0.94, σ1 = 0.03, σ2 = 0.19, μ1 = 0.08, and μ2 = 0.44 (see Table S5 for standard errors). A Pearson χ2 test indicated that the optimized distribution fits the data well (the fitted model did not differ significantly from the data; p = 0.231 > 0.05). The distribution is dominated by fibers and fragments (CSF = 0.25–0.75) but also has a distinct second peak at a CSF of 0.07, which is mainly attributed to sheets (Figure 2A). The distribution captures the main features of shapes encountered in the environment as well as the relative abundances of these (now continuous) shapes in one go. Most illustrative though is the continuous character of microplastic shape.

$y = f1(\frac{1}{sqrt(2pi()\sigma_1^2})e^{-(x-\mu_1)^2/2\sigma_1^2} + f2(\frac{1}{sqrt(2pi()\sigma_2^2})e^{-(x-\mu_2)^2/2\sigma_2^2}$

For polymer abundance, for the K&K2019 paper I used the fixed values from Burns & Boxall review, and went straight for density. Given these fixed relative abundances, you could calculate polymer type randomly for N particles via: 

pols <- c("PE", "PP", "PS")
rel.ab <- c(0.5, 0.3, 0.2)
n = 1000
sample(pols, size = n, prob = rel.ab, replace = TRUE)

Particle volume is determined according to shape as follows:

$V = \frac{\pi}{6}L^3CSF^2$

Where CSF = corey shape factor and L = length.

Particle mass is estimated as follows:

$m = pV*\frac{1}{1e12}*1000$

Where *m* is the mass (mg), *p* is density (g/cm^3), *V* is volume (um^3) - which is calculated by the cube of the length (um) of each particle, and additional conversion factors for mg to g (x1000) and cm^3 to um^3 (1e-12).
```{r, include=FALSE}
#variables to set for funcions
      xmin = 1 #UM
      alpha = 1.6 #2.7 is preferred in unpublished Kooi et al.
      
      mu1 <- 0.08
      mu2 <- 0.44
      sd1 <- 0.03
      sd2 <- 0.19
      lambda1 <- 0.06
      lambda2 <- 0.94
        
      d.alpha = 73.8 #tail heaviness
      d.beta = 69.9  #asymmetry
      d.mu = 0.840   #location
      d.delta = 0.0972 #scale
X.func <- function (X){
  success <- FALSE
  while (!success){
    U = runif(1, 0, 1)
    X = xmin*(1-U)^(1/(1-alpha))
    success <- X < 5000} ##should be smaller than 5000 um 
  return(X)
}
D.func <- function (D){
  success <- FALSE
  while (!success){
    D = rnig(1, mu = d.mu, alpha = d.alpha, beta = d.beta, delta = d.delta)
    success <- D < 2.63} ## include upper limit of 2.63, the max. 
  return(D)
}
synthetic_distributed_data_builder <- function(count, volume_l, addedfactors){
  #Preset parameters for pdfs
   ## Generate values for the three distributions
      set.seed(123)
      Data <- data.frame(Size = numeric(0))
      
      for(i in 1:count){
        X <- X.func()
        Data <- rbind(Data, X)
      }
      
      ##++++++++++++++++++++++++++++++++++++++++
      ## SIZE DISTRIBUTION
      ##++++++++++++++++++++++++++++++++++++++++
      
      colnames(Data) <- c("Size")
      
      #min(Data$Size) ##20 um
      #max(Data$Size) ##5000 um
      
      ##++++++++++++++++++++++++++++++++++++++++
      ## SHAPE DISTRIBUTION
      ##++++++++++++++++++++++++++++++++++++++++
      
      #Sample N random uniforms U
      U =runif(count)
      
      #Sampling from the mixture
      for(i in 1:count){
        if(U[i]<lambda1){
          Data$Shape[i] = rtnorm(1,mu1,sd1, lower = 0, upper = 1)
        }else{
          Data$Shape[i] = rtnorm(1,mu2,sd2, lower = 0, upper = 1)
        }
      }
      
      min(Data$Shape) ##0
      max(Data$Shape) ##1
      
      ##++++++++++++++++++++++++++++++++++++++++
      ## DENSITY DISTRIBUTION
      ##++++++++++++++++++++++++++++++++++++++++
      
      Dens <- data.frame(Density = numeric(0));
      
      for(i in 1:count){
        X <- D.func()
        Dens <- rbind(Dens, X)
      }
      
      colnames(Dens) <- c("Density")
      
      Data <- cbind(Data, Dens)
      
      ##++++++++++++++++++++++++++++++++++++++++
      ## Polymer DISTRIBUTION
      ##++++++++++++++++++++++++++++++++++++++++
      
      pols <- c("PE", "PP", "PS")
      rel.ab <- c(0.5, 0.3, 0.2)
      polymer <- sample(pols, size = count, prob = rel.ab, replace = TRUE)
      Data <- cbind(Data, polymer)
      
      Data %>%  
        mutate(size.category = factor(case_when(
          Size < 10 & Size >= 1 ~ "1µm < 10µm",
          Size < 100 & Size >= 10 ~ "10µm < 100µm",
          Size < 1000 & Size >= 100 ~ "100µm < 1mm",
          Size < 5000 & Size >= 1000 ~ "1mm < 5mm"))) %>% 
        mutate(mass.mg = Density * Size^3 * 1E-9) %>% 
        mutate(um3 = (Size^3)) %>%
        bind_cols(addedfactors[rep(1, times = nrow(.)),]) %>% 
        mutate(vol.L = volume_l) %>% 
        mutate(vol.mL = vol.L * 1000) %>% 
        mutate(dose.um3.mL.master = um3 / vol.mL) %>% 
        mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
        mutate(dose.mg.L.master = mass.mg / vol.L) %>% 
        mutate(particles.total = factor(as.character(count)))
}
add_prediction <- function(fun, model, ...){
        fun(...) %>%
        mutate_if(~is.numeric(.) && (.) > 0, log10) %>%
        bind_cols(predict(model, newdata = ., type = "prob") %>% 
        rename("ToxicityProbability" = `1`))                  
}
```


Test synthetic data.

```{r}
##++++++++++++++++++++++++++++++++++++++++
## Quick check results
##++++++++++++++++++++++++++++++++++++++++
skim(synthetic_data_builder(count = 10000, 
                                    volume_l = 1000, 
                                    addedfactors = possible_variables[488,]
                                    ))
```


### Uniform Distribution

It may be possible to visualize discrete relationships of toxicity by creating a uniform distribution of particles within the parameter space of the training dataset, then plotting the predicted toxicities for different predictors.

#### Generate uniform particles
```{r}
n = 1000
#Create uniform characteristics for base particle characteristics (size, shape, density)
csf <- runif(n, min = min(aoc_z$csf, na.rm = T), max(aoc_z$csf, na.rm = T)) #Set this to parameter space 
#mass.per.particle.mg <- runif(n, min = 0, max = 1.65e+1)
density.mg.um.3 <- runif(n, min = min(aoc_z$density.mg.um.3, na.rm = T), max = max(aoc_z$density.mg.um.3, na.rm = T))
size.length.um.used.for.conversions <- runif(n, min = min(aoc_z$size.length.um.used.for.conversions, na.rm = T), max = max(aoc_z$size.length.um.used.for.conversions, na.rm = T))
#create uniform distribution of polymer types
#pols <- c("PE", "PP", "PS", "PVC", "PET")
#rel.ab <- c(0.2, 0.2, 0.2, 0.2, 0.2)
#polymer <- sample(pols, size = n, prob = rel.ab, replace = TRUE)
#create data frame with uniform distirbutions
unif_df <- data.frame(csf, density.mg.um.3, size.length.um.used.for.conversions) %>%
  mutate(mass.per.particle.mg = density.mg.um.3 * size.length.um.used.for.conversions ^ 3 * 1E-9) %>% 
  mutate(particle.volume.um3 = pi/6 * (size.length.um.used.for.conversions ^ 3) * csf ^2 ) %>% #equation 4
  #volume
  mutate(vol.mL = 0.0001) %>% 
  mutate(vol.L = vol.mL / 1000) %>%  
  mutate(dose.um3.mL.master = particle.volume.um3 / vol.mL) %>% 
  mutate(dose.particles.mL.master = 1 / vol.mL) %>% 
  mutate(dose.mg.L.master = mass.per.particle.mg / vol.L) %>% 
    #organism/test system characteristics
  mutate_if(is.character, as.factor)
skim(unif_df)
```
HIGHLIGHT AREA UNDER THE CURVE
If we go down to lower number, we would not adequately mimic distribution. Stick to creating 1E5 particles, then change volume. 
Could capture variability for small bumbers by simulating small number of particles


How to calculate uncertainties? 
  Replace effect parameters by median +- SD, then predict tox for each distribution, plot, then if it is minor, FORGET about it! REsiduals in Merel's models were not randomly distributed -  so imperfect models.
  
What sizes should be used?
  1-5,000 um. 

Wait for Merel to update interactions between shape/size/density/count.
  
What about max size ingest?
  Tox studies typically do not exposure animals to particles bigger than their mouth size opening, so model may not understand that relationship. 
  Could impose bioavailable fraction AFTER model. JUST do predictions for species that have bioavailable fraction data. Could estimate mouth size from prey to predator fraction. JAMS et al has mouth-size opening parameters.

Non-linear regression to get dose-response. 

Mattson, Hasselof, Frontiers in Science - measures size distribution. 

Final figure: SINGLE toxicity prediction for each dose. Use environmental data. Make SSD. Risk Characterization. The model's SSD is fundamentally different from the SSD made from monodisperse particles, so it is MORE reliable. 
Big assumption: all particle toxicities are ADDITIVE! Depends on effect mechanism - food dilution. 


# Dose-response curves
#Part of me just wants to set the uniform distribution to the distinct version of the dataset. That way we can define uncertainty for each dose response curve and are not extrapolating within ranges (sometimes people just get data for the max and min in a distribution that doesn't mean that we know what the middle should be.)

```{r}
org_endpoint_comb <- aoc_z %>%
  dplyr::select(organism.group, bio.org, lvl1_f) %>%
  distinct() 
df_expand <- org_endpoint_comb %>%
  expand_grid(unif_df)
#Ran the code below too to see if the glm methods were the same. They are returning the same response. 
df_expand$prob <- predict(final_model, df_expand, type = "response")
for(row in 1:nrow(org_endpoint_comb)){
  slice <- org_endpoint_comb[row,]
  doses <- inner_join(df_expand, slice) %>%
    dplyr::select(-organism.group, -bio.org, -lvl1_f) %>%
    pivot_longer(cols = -prob, names_to = "type", values_to = "values")
 plot <- ggplot(doses) + geom_point(aes(x = values, y = prob)) + facet_wrap(type~., scales = "free") + labs(title = paste0(as.character(unlist(slice)), collapse = "_"))
  ggsave(plot = plot, filename = paste0(paste0(as.character(unlist(slice)), collapse = "_"), ".png"), path = "Tox Data/figures")
}  
#Testing this out for actual data, not extrapolated uniform data using the training dataset.
org_endpoint_comb <- actual_model_data %>%
  dplyr::select_if(function(col) is.character(col) | 
                                   is.factor(col)) %>%
  distinct() 
df_expand <- actual_model_data 
for(row in 1:nrow(org_endpoint_comb)){
  slice <- org_endpoint_comb[row,]
  doses <- inner_join(df_expand, slice) %>%
    dplyr::select_if(is.numeric) %>%
    pivot_longer(cols = -.outcome, names_to = "type", values_to = "values")
 plot <- ggplot(doses) + geom_point(aes(x = values, y = .outcome)) + facet_wrap(type~., scales = "free") + labs(title = paste0(as.character(unlist(slice)), collapse = "_"))
  ggsave(plot = plot, filename = paste0(paste0(as.character(unlist(slice)), collapse = "_"), ".png"), path = "Tox Data/figures")
}  
```

# SSD

```{r}
df_expand_sensitivity_distribution <- df_expand %>%
  filter(prob > 0.5) %>%
  group_by(organism.group, bio.org, lvl1_f) %>% #Can add other factors here if we want to flesh this out for each organism.
  summarise(dose.um3.mL.master = min(dose.um3.mL.master), dose.surface.area.um2.mL.master = min(dose.surface.area.um2.mL.master)) %>%
  ungroup()
ggplot(df_expand_sensitivity_distribution) + stat_ecdf(aes(x = dose.surface.area.um2.mL.master))
ggplot(df_expand_sensitivity_distribution) + stat_ecdf(aes(x = dose.um3.mL.master))
#Other Option with automated extraction from the dataset. 
df_expand_sensitivity_distribution <- df_expand %>%
  filter(.outcome > 0.5) %>%
  group_by_if(function(col) is.character(col) | 
                                   is.factor(col)) %>% #Can add other factors here if we want to flesh this out for each organism.
  summarise_if(is.numeric, min) %>%
  ungroup()
ggplot(df_expand_sensitivity_distribution) + stat_ecdf(aes(x = dose.particles.mL.master))
```

#Run Merels distribution through the model to predict the liklihood of risk in the environment for organism endpoint combinations. 


# Sensitivity Analysis
it would be great to use the full approach to find out which model parameters and MP subclasses the HC5 is most sensitive to.And also if that HC5 is sensitive to which of Merel’s compartment-specific MP parameterisations is used.

# Shiny App Tabs
1) ML predictor
2) Chemical data
  Bioaccumulation model, partitioning with chemicals/particles. Would add toxicity from chemicals to toxicity from particles! Use same model from Worm paper. Would show how it would compare with toxicities from particles. We have criticial body burden concept (lipids) 
© 2021 GitHub, Inc.