---
title: "Global Microplastics Risk Characterization"
author: "Scott Coffin"
date: "3/10/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r}
library(tidyverse)
library(calecopal)
library(ssdtools)
library(DT)
library(plotly)
library(gridExtra)
library(grid)
library(wesanderson)
library(ggdark)
library(broom)
library(knitr)
library(ggdark)
library(viridis)
library(ggpubr)
#library(overlapping)
```

# Hazard Concentration and Alignment Parameters
```{r}
###### Choose concentration alignment parameters####
alpha = 2.07 #personal comm. with Bart suggests 2.07 for marine surface water #(use 1.6 or 2.5)
x2D_set = 5000 #100 #upper size range (microns)
x1D_set = 1 #1 #lower size range (microns)

#### Choose assessment factor ####
AF = 1#5

##### Choose hazard concentration and confidence intervals####
#particles/L
Threshold1 = 0.5 / AF
Threshold2 = 8.1 / AF
Threshold3 = 14.1 / AF
Threshold4 = 93.8 / AF
```

```{r Theme, include=FALSE}
#Theme type
     theme.type <- theme_bw(base_size = 15) +
                       #dark_theme_bw(base_size = 18)
   theme(plot.title = element_text(hjust = 0.5),
                    plot.subtitle = element_text(hjust = 0.5))
     #color selection
     fill.type <-    scale_fill_viridis(discrete = TRUE)#,
                         #scale_fill_brewer(palette = "Paired"),
                         # scale_fill_tron(),
                         # scale_fill_locuszoom(),
                         # scale_fill_d3(),
                         # scale_fill_npg(),
                         # scale_fill_jama())
     #color selection
     color.type <- scale_color_viridis(discrete = TRUE)#,
                         # scale_color_brewer(palette = "Paired"),
                         # scale_color_tron(),
                         # scale_color_locuszoom(),
                         # scale_color_d3(),
                         # scale_color_npg(),
                         # scale_color_jama())

```



```{r}
# print table
kable(data.frame(
  Category = c("Hazard Concentrations (particles/L)"),
  "Threshold1" = Threshold1,
  "Threshold 2" = Threshold2,
  "Threshold 3" = Threshold3,
  "Threshold 4" = Threshold4))
```

```{r}
#print table of sizes
kable(data.frame(
  Category = c("Concentration Alignment Parameters"),
  alpha = alpha,
  "lower_size_range_microns" = x1D_set,
  "upper_size_range_microns" = x2D_set))
```
To compare thresholds to concentrations observed in the environment, datasets compiled in previous meta-analyses were collated (Covernton et al. 2019; Adam, Yang, and Nowack 2019).

Adam, Véronique, Tong Yang, and Bernd Nowack. 2019. “Toward an Ecotoxicological Risk Assessment of Microplastics: Comparison of Available Hazard and Exposure Data in Freshwaters.” Environmental Toxicology and Chemistry 38 (2): 436–47. https://doi.org/10.1002/etc.4323.

Covernton, Garth A., Christopher M. Pearce, Helen J. Gurney-Smith, Stephen G. Chastain, Peter S. Ross, John F. Dower, and Sarah E. Dudas. 2019. “Size and Shape Matter: A Preliminary Analysis of Microplastic Sampling Technique in Seawater Studies with Implications for Ecological Risk Assessment.” Science of The Total Environment 667 (June): 124–32. https://doi.org/10.1016/j.scitotenv.2019.02.346.



```{r}
#data import
# Adam et al (2019) data needs correcting
adam <- read.csv("Concentration Data/Datasets/adam2019.csv", na.strings = "N.A.") %>% 
  rename(Sampling.apparatus = ï..Sampling.apparatus) %>% 
  mutate(x1M = Min.particle.size.um,
         x2M = Max.particle.size.um) %>% 
  mutate(Sample.Type = "sample")

#SF bay data from Zhu et al (2021)
sfBay <- read.csv("Concentration Data/Datasets/SFBayData.csv", na.strings = "NA", stringsAsFactors = TRUE) %>%
  rename(sampleID = ï..sampleID,
          x1M = Min.particle.size.um,
         x2M = Max.particle.size.um) %>% 
  mutate(Sample.Type = "sample")
# import from Covernton et al (2019)
covernton <- read.csv("Concentration Data/Datasets/Covernton.csv", na.strings = "NA", stringsAsFactors = TRUE) %>%
  rename(x1M = Min.particle.size.um,
         x2M = Max.particle.size.um) %>% 
  mutate(Sample.Type = "sample")
```


## Size-based Correction
Given an upper limit (UL) and lower limit (LL) of the measured and default size range, a dimensionless correction factor ($CF_{meas}$) for measured environmental concentrations may be calculated, which rescales the measured number concentrations for a certain size range to the number concentration for the microplastics default (D) size range (e.g. 1 to 5,000 um) (Koelmans et al 2020). Coefficients for the environmental-specific compartment for the power law distribution for length (L) with slope $\alpha_L$ is from Table S4 of Kooi et al (2021). 

$CF_{Meas} = \frac{L^{1-a}_{UL,D} - L^{1-a}_{LL,D}}{L^{1-a}_{UL,M} - L^{1-a}_{LL,M}}$

```{r}
# Align Data

#function to derive correction factor (CF) from Koelmans et al (equation 2)
CFfnx = function(a = alpha, #default alpha from Koelmans et al (2020)
                 x2D = x2D_set, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
                 x1D = x1D_set, #1 um is lower default size
                 x2M, x1M){
  
  CF = (x2D^(1-a)-x1D^(1-a))/(x2M^(1-a)-x1M^(1-a))
  
  return(CF)
}
#verify it works (expected answer is 40.37)
#CFfnx(x1M = 333, x2M = 5000)

adam <- adam %>% 
  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
  mutate(particles.m3.corrected = CF * Single.Measurement.conc....m3.) %>% #convert single concenetrations
  mutate(particles.m3.corrected_mean = CF * Mean.conc....m3.) %>%  #convert mean concentrations from distributions
  mutate(particles.m3.corrected_median = CF * Median.conc....m3.) %>%   #convert mean concentrations from distributions
  mutate(particles.single.median.m3 = ifelse(is.na(particles.m3.corrected), particles.m3.corrected_median, particles.m3.corrected)) %>% 
  mutate(particles.m3.master = ifelse(is.na(particles.single.median.m3), particles.m3.corrected_mean, particles.single.median.m3)) %>% 
  mutate(particle.L.master = particles.m3.master/1000) %>% 
  filter(particle.L.master > 0) %>% 
  mutate(System = factor(System)) %>% 
  mutate(source = "Adam")

# correct aquatic concentrations for SF Bay
sfBay <- sfBay %>% 
  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
  mutate(particle.L.master = particles.L.blank.corrected * CF) %>% 
  mutate(source = "SFEI") %>% 
  mutate(Reference = "Zhu et. al 2021") %>% 
  rename(Sampling.location = general.location) %>% 
  select(-c(particles.L.blank.corrected, particles.kg.blank.corrected, particles.fish.blank.corrected))

# correct aquatic concentrations for Ocean Samples
covernton <- covernton %>% 
  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
  mutate(particle.L.master = conc....m3. * CF) %>% 
  mutate(source = "Covernton") %>% 
  rename(Reference = study) %>% 
  rename(Sampling.location = Sampling.Location) %>% 
  select(c(Sampling.apparatus, Sampling.location, x1M, x2M, Sample.Type, System, CF,particle.L.master, source, Reference)) %>% 
   mutate(sampleID = NA,
         specific.location = NA,
         latitude = NA,
         longitude = NA,
         sample.matrix = "water",
         sample.matrix.specific = "surface water"
         )

## join datasets ##
#first ensure that both datasets share all columns
adam_sub <- adam %>% 
  select(c(Sampling.apparatus, Sampling.location, x1M, x2M, Sample.Type, System, CF,particle.L.master, source, Reference)) %>% 
  mutate(sampleID = NA,
         specific.location = NA,
         latitude = NA,
         longitude = NA,
         sample.matrix = "water",
         sample.matrix.specific = "surface water"
         )

#join data
globalData <- rbind(adam_sub, sfBay, covernton) %>% 
  mutate(Conc = particle.L.master) %>% 
  mutate_if(is.character, as.factor)
```

## Global Exceedances
```{r}
#make new dataframe to plot both histograms together
sampleSimpleADAM <- globalData %>%
  filter(sample.matrix.specific == "surface water") %>% 
  filter(System == "Ocean") %>% 
  drop_na(Conc) %>% 
  droplevels()

#make new dataframe to plot both histograms together
dfADAM <- rbind(sampleSimpleADAM)#,food.dilution.simple)

#calculate exceedance
dfADAM_exceedance <- dfADAM %>% 
  mutate(aboveThreshold = factor(case_when(
    Conc < Threshold1 ~ "below Threshold 1",
    Conc >= Threshold1 & Conc < Threshold2 ~ "above Threshold 1",
    Conc >= Threshold2 & Conc < Threshold3 ~ "above Threshold 2",
    Conc >= Threshold3 & Conc < Threshold4 ~ "above Threshold 3",
    Conc >= Threshold4 ~ "above Threshold 4"
  )))

#give summary stat for exceedance
exceedance <- dfADAM_exceedance  %>%
  filter(Sample.Type == "sample") %>% 
  dplyr::select(c(Conc, aboveThreshold)) %>%
  group_by(aboveThreshold) %>%
  dplyr::summarize(n = n()) %>% 
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%"))

kable(exceedance)
```
### Concentrations Histogram
```{r}
binwidth <- 0.25

globalData %>% 
  ggplot(aes(x = particle.L.master,
             fill = source,
            # color = source, 
             binwidth = binwidth)) +
  # geom_histogram(aes(y = ..count.. / (sum(..count..) * binwidth)), 
  #                binwidth = binwidth,
  #                alpha = 0.9,
  #                position = "stack") +
  geom_density( alpha = 0.6, lwd = 0.8, adjust = 0.6) +
  scale_fill_manual(name = "Location", labels = c("Rivers and Lakes",
                                                  "Oceans",
                                                  "San Francisco Bay"
                                                  ),
                    values = cal_palette("superbloom3")) + 
                   #values = c("#00798c", "#d1495b", "black")) +
  scale_x_continuous(name = "Particles/L",breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif, trans = "log10")+
  ylab("Normalized Abundance") +
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  labs(title = "Global Surface Water Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("n = 70 studies, 95 sampling locations, n_samples = 747; corrected for size, blanks. Alpha = ",alpha))+
  dark_theme_minimal(base_size = 25) +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(size = 12))
```
### Concentrations boxplot
```{r}
binwidth <- 0.25

globalData %>% 
  ggplot(aes(x = particle.L.master,
             fill = source,
            # color = source, 
             binwidth = binwidth)) +
 geom_boxplot() +
   scale_fill_manual(name = "Location", labels = c("Rivers and Lakes",
                                                  "Oceans",
                                                  "San Francisco Bay"
                                                  ),
                    values = cal_palette("superbloom3")) + 
                    #values = c("#00798c", "#d1495b")) +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif, trans = "log10")+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  labs(title = "Global Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("n = 70 studies, 95 sampling locations, n_samples = 747; corrected for size, blanks. Alpha = ",alpha))+
  theme.type +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

### Risk Characterization Histogram (Global Marine Data)
```{r}
#generate plot
dfADAM_exceedance %>% 
  filter(Conc > 0) %>% 
  filter(Sample.Type == "sample") %>%
  filter(System %in% c("Ocean", "Estuary")) %>% 
  ggplot(aes(x = Conc, fill = aboveThreshold))+
  geom_histogram(aes(y = ..count../sum(..count..)),bins = 50, alpha = 0.9, position = "identity") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.045), label = paste(Threshold1,"particles/L"),  color = "goldenrod2") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.055), label = ("Threshold1"),  color = "goldenrod2") +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'goldenrod2', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'tomato1', size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'mediumvioletred', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'gray33', size = 1.3) +
  #geom_text(label = "Threshold 1", color = 'goldenrod2', x = log10(Threshold1) - 0.8*log10(Threshold1), y = 0.055, size = 6)+
#   geom_text(label = "Threshold 2", color = 'tomato1', x = log10(Threshold2) - 0.8*log10(Threshold2), y = 0.055, size = 6)+
 #  geom_text(label = "Threshold 3", color = 'mediumvioletred', x = log10(Threshold3) - 0.8*log10(Threshold3), y = 0.055, size = 6)+
  # geom_text(label = "Threshold 4", color = 'gray33', x = Threshold4 - 0.8*log10(Threshold4), y = 0.055, size = 6)+
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif, trans = "log10")+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  #coord_cartesian(xlim = c(0,100000000)) +
  #scale_x_continuous(labels = scales::scientific) +
   xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)")) +
  scale_y_continuous(name = "Relative Density", labels = scales::percent)+
  scale_fill_manual(values = c("below Threshold 1"= "cyan3", 
                               "above Threshold 1" = "goldenrod2", 
                               "above Threshold 2" = "tomato1",
                               "above Threshold 3" = "violetred4", 
                               "above Threshold 4" = "gray33")) +
  labs(title = "Global Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from n = 70 studies, 95 sampling locations, n_samples = 747; corrected for size, blanks. Alpha = ",alpha))+
  theme.type +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

### SFBay Exceedances
```{r}
#make new dataframe to plot both histograms together
sampleSimpleSFEI <- globalData %>%
  filter(source == "SFEI") %>% 
  filter(sample.matrix.specific == "surface water") %>% 
  drop_na(Conc) %>% 
  droplevels()

#make new dataframe to plot both histograms together
dfSFEI <- rbind(sampleSimpleSFEI)#,food.dilution.simple)

#calculate exceedance
dfSFEI_exceedance <- dfSFEI %>% 
  mutate(aboveThreshold = factor(case_when(
    Conc < Threshold1 ~ "below Threshold 1",
    Conc >= Threshold1 & Conc < Threshold2 ~ "above Threshold 1",
    Conc >= Threshold2 & Conc < Threshold3 ~ "above Threshold 2",
    Conc >= Threshold3 & Conc < Threshold4 ~ "above Threshold 3",
    Conc >= Threshold4 ~ "above Threshold 4"
  )))

#give summary stat for exceedance
exceedance <- dfSFEI_exceedance  %>%
  filter(Sample.Type == "sample") %>% 
  dplyr::select(c(Conc, aboveThreshold)) %>%
  group_by(aboveThreshold) %>%
  dplyr::summarize(n = n()) %>% 
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%"))

kable(exceedance)
```

### Risk Characterization Histogram (SFBay)
```{r}
#generate plot
dfSFEI_exceedance %>% 
 # filter(Conc > 0) %>% 
  filter(Sample.Type == "sample") %>% 
  ggplot(aes(x = Conc, fill = aboveThreshold))+
  geom_histogram(aes(y = ..count../sum(..count..)),bins = 38, alpha = 0.9, position = "identity") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.045), label = paste(Threshold1,"particles/L"),  color = "goldenrod2") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.055), label = ("Threshold1"),  color = "goldenrod2") +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'goldenrod2', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'tomato1', size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'mediumvioletred', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'gray33', size = 1.3) +
  #geom_text(label = "Threshold 1", color = 'goldenrod2', x = log10(Threshold1) - 0.8*log10(Threshold1), y = 0.055, size = 6)+
#   geom_text(label = "Threshold 2", color = 'tomato1', x = log10(Threshold2) - 0.8*log10(Threshold2), y = 0.055, size = 6)+
 #  geom_text(label = "Threshold 3", color = 'mediumvioletred', x = log10(Threshold3) - 0.8*log10(Threshold3), y = 0.055, size = 6)+
  # geom_text(label = "Threshold 4", color = 'gray33', x = Threshold4 - 0.8*log10(Threshold4), y = 0.055, size = 6)+
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif, trans = "log10")+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  #coord_cartesian(xlim = c(0,100000000)) +
  #scale_x_continuous(labels = scales::scientific) +
   xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)")) +
  scale_y_continuous(name = "Relative Density", labels = scales::percent)+
  scale_fill_manual(values = c("below Threshold 1"= "cyan3", 
                               "above Threshold 1" = "goldenrod2", 
                               "above Threshold 2" = "tomato1",
                               "above Threshold 3" = "violetred4", 
                               "above Threshold 4" = "gray33")) +
  labs(title = "San Francisco Bay Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("n = 107; corrected for size, blanks. Alpha = ",alpha))+
  theme.type +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```
### ECDF (SF Bay)

```{r}
#ECDF by System
globalData %>% 
  filter(source == "SFEI") %>% 
  drop_na(particle.L.master) %>% 
  filter(particle.L.master > 0.0) %>% 
filter(System %in% c("Ocean", "Estuary")) %>% 
  ggplot(aes(x = particle.L.master)) + #, color = System))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.5, color = "black") +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.3, size = 1, color = "black") + #'azure3') +
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "Threshold 1", color = "#00798c" , x = Threshold1, y = 0.15, size = 6)+
  geom_text(label = "Threshold 2", color = "#66a182" , x = Threshold2, y = 0.22, size = 6)+
  geom_text(label = "Threshold 3", color = '#edae49', x = Threshold3, y = 0.35, size = 6)+
  geom_text(label = "Threshold 4", color = "#d1495b", x = Threshold4, y = 0.45, size = 6)+
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
  scale_y_continuous(labels = scales::percent)+
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  #annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  theme.type #+
  #labs(title = "Global Surface Water Marine Microplastics Concentrations",
   #    subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
    #   caption = paste("Concentration data from Adams et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))
 
```


### ECDF (Global)
```{r}
#ECDF by System
globalData %>% 
  drop_na(particle.L.master) %>% 
  filter(particle.L.master > 0.0) %>% 
filter(System %in% c("Ocean", "Estuary")) %>% 
  ggplot(aes(x = particle.L.master)) + #, color = System))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.5, color = "black") +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.3, size = 1, color = "black") + #'azure3') +
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "Threshold 1", color = "#00798c" , x = Threshold1, y = 0.15, size = 6)+
  geom_text(label = "Threshold 2", color = "#66a182" , x = Threshold2, y = 0.22, size = 6)+
  geom_text(label = "Threshold 3", color = '#edae49', x = Threshold3, y = 0.35, size = 6)+
  geom_text(label = "Threshold 4", color = "#d1495b", x = Threshold4, y = 0.45, size = 6)+
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
  scale_y_continuous(labels = scales::percent)+
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  #annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  theme.type #+
  #labs(title = "Global Surface Water Marine Microplastics Concentrations",
   #    subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
    #   caption = paste("Concentration data from Adams et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))
 
```

### ECDF (SFEI by Matrix)
```{r}
globalData %>% 
  filter(source == "SFEI") %>% 
  filter(sample.matrix == "water") %>% 
  drop_na(particle.L.master) %>% 
  ggplot(aes(x = log10(particle.L.master), color = sample.matrix.specific))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.7) +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.5, size = 1.5) +
  scale_color_manual(name = "Matrix",values = wes_palette("Darjeeling1"))+
  # geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1.3) +
  # geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1.3) +
  # geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1.3) +
  # geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1.3) +
  # geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
  # geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
  # geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
  # geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
  scale_y_continuous(labels = scales::percent)+
  #coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, " um)"))+
  theme.type +
  labs(title = "San Francisco Bay Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from Zhu et al (2021): 29 sampling locations, n = 136; corrected for size ranges. Alpha = ",alpha))+
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

### ECDF (SFEI by Mesh Size)
```{r}
globalData %>% 
  filter(source == "SFEI") %>% 
  filter(sample.matrix == "water") %>% 
  drop_na(particle.L.master) %>% 
  ggplot(aes(x = log10(particle.L.master), color = Sampling.apparatus))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.7) +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.5, size = 1.5) +
  scale_color_manual(name = "Matrix",values = wes_palette("Darjeeling1"))+
  # geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1.3) +
  # geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1.3) +
  # geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1.3) +
  # geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1.3) +
  # geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
  # geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
  # geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
  # geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
  scale_y_continuous(labels = scales::percent)+
  #coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, " um)"))+
  theme.type +
  labs(title = "San Francisco Bay Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from Zhu et al (2021): 29 sampling locations, n = 136; corrected for size ranges. Alpha = ",alpha))+
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

### Boxplot SFEI
```{r}
globalData %>% 
  filter(source == "SFEI") %>% 
  filter(sample.matrix == "water") %>% 
  drop_na(particle.L.master) %>% 
  ggplot(aes(y = particle.L.master, x = sample.matrix.specific))+
  geom_boxplot(alpha = 0.4) +
  geom_point(aes(color = Sampling.apparatus)) +
  scale_color_manual(name = "Sampling Apparatus",values = wes_palette("Darjeeling1"))+
   # geom_hline(yintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1) +
   # geom_hline(yintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1) +
   # geom_hline(yintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1) +
   # geom_hline(yintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1) +
  # geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
  # geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
  # geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
  # geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
  scale_y_log10(breaks = scales::trans_breaks("log10", function(y) 10^y),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  ylab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, " um)"))+
  theme.type +
  labs(title = "San Francisco Bay Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from Zhu et al (2021): 29 sampling locations, n = 136; corrected for size ranges. Alpha = ",alpha))+
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

## Modelling

```{r eval=FALSE, include=FALSE}
surfaceData <- globalData %>% 
  filter(sample.matrix.specific == "surface water") %>% 
  dplyr::select(Conc) %>% 
  drop_na(Conc)

sample_dists_ADAM <- ssd_fit_dists(surfaceData, #data frame
                              left = "Conc", #string of the column in data with the concentrations
                              # right = left, #string of the column with the right concentration values. If different from left, then the data are considerd to be censored
                              dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), #char vector of distribution anmes
                              computable = FALSE, #flag specifying whether to only return fits with numerically computable standard errors
                              silent = FALSE) #flag indicating whether fits should fail silently

autoplotADAM<- autoplot(sample_dists_ADAM) #plots the distribution in ggplot2
autoplotADAM
ssd_gof(sample_dists_ADAM) #check the goodness of fit
#there are multiple fitting distributions, so check which fits best
sample_gof_ADAM <- ssd_gof(sample_dists_ADAM)
sample_gof_ADAM[order(sample_gof_ADAM$delta), ] #orders by delta. Use the aicc (Akaike's Information Criterion corrected for sample size) for model selection 
write.csv(sample_gof_ADAM,"Concentration data/sample_gof_ADAM.csv")
#choose the distribution that you want to plot
sample_dists_ADAM_choice <- ssd_fit_dists(samplesADAM, #data frame
                                   left = "Conc", #string of the column in data with the concentrations
                                   # right = left, #string of the column with the right concentration values. If different from left, then the data are considerd to be censored
                                   dists = c("lgumbel"), #char vector of distribution anmes
                                   computable = FALSE, #flag specifying whether to only return fits with numerically computable standard errors
                                   silent = FALSE) #flag indicating whether fits should fail silently
set.seed(99)
sample_pred_ADAM <- predict(sample_dists_ADAM_choice,
                       average = FALSE,
                       ic = "aicc",
                       nboot = 10,
                       ci= TRUE) #estimates model-averaged estimates based on aicc

sample_pred_ADAM # The resultant object is a data frame of the estimated concentration (est) with standard error (se) and lower (lcl) and upper (ucl) 95% confidence limits by percent of species affected (percent). The confidence limits are estimated using parametric bootstrapping.

```

```{r eval=FALSE, include=FALSE}
sample_pred_ADAM %>% mutate_if(is.numeric, ~ signif(., 3)) %>% 
  datatable(rownames = FALSE,
            extensions = c('Buttons', 'Scroller'),
            options = list(
              dom = 'Brftp',
              scrollY = 400,
              scroller = TRUE,
              buttons = c('copy', 'csv', 'excel')), 
            class = "compact",
            colnames = c("Percent", "Estimated Mean Concentration", "Standard Error", "Lower 95% Confidence Limit", "Upper 95% Confidence Limit", "Distribution"),
            caption = "Predicted Concentration distribution with uncertanties."
  )

#order data
samplesADAM <- samplesADAM %>% 
  filter(System != "") #take out blanks

sampleSSDADAM <- samplesADAM[order(samplesADAM$Conc), ]
sampleSSDADAM$frac <- ppoints(samplesADAM$Conc, 0.5)
```


```{r eval=FALSE, include=FALSE}
ECDF_model_occurrence_ADAM <- ggplot(sample_pred_ADAM,aes_string(x = "est")) +
  geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "#81a88d", fill = "#81a88d") +
  geom_line(aes_string(y = "percent/100"), linetype = 'dashed', alpha = 0.8) +
  geom_point(data = sampleSSDADAM,aes(x = Conc, y =frac, color = System), size =1) + 
  #geom_text(data = sampleSSD, aes(x = Conc, y = frac, label = Location), hjust = 1.1, size = 4) + #season labels
  scale_y_continuous("Cumulative Distribution (%)", labels = scales::percent) +
  #expand_limits(y = c(0, 1)) +
  xlab("Concentration (particles/L)")+
  labs(title = "Adam et al 2019 Microplastics Concentration Cumulative Distribution Function",
       subtitle = "Smoothing/95% CI ribbon based on average of log-logical and log-normal Distributions Fit",
       caption = "Adam et al 2019 data; sampling corrected to 1-5,000 um") +
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  scale_color_manual(values = wes_palette("Darjeeling2"))

#white mode
ECDF_model_occurrence_ADAM_white <- ECDF_model_occurrence_ADAM +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 'red') +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 	'red') +
  geom_text(label = "5% HC: 95% LCL", color = 'red', x = Threshold3, y = 0)+
  geom_text(label = "5% hazard concentration", color = 'red', x = 110, y = 0.03)+
  geom_text(label = "5% HC: 95% UCL", color = 'red', x = Threshold2, y = 0)+
  geom_text(x = 110, y = 0, label = paste(Threshold1,"particles/L"), color = 'red') +  #label for hazard conc
  #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "#A2A475") +
  #geom_text(label = "92.5% samples below 5% HC Mean", x = 4.5, y = 0.94, color = "#A2A475") +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
ECDF_model_occurrence_ADAM_white

```

```{r eval=FALSE, include=FALSE}
#dark mode
ECDF_model_occurrence_ADAM_dark <- ECDF_model_occurrence_ADAM +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'goldenrod2', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'tomato1', size = 1.3) +
 geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'mediumvioletred', size = 1.3) +
 geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'gray33', size = 1.3) +
  #geom_text(x = Threshold1, y = 0.1, label = paste(Threshold1,"particles/L"), color = 'red', size =5) +  #label for hazard conc
  #geom_hline(yintercept = 0.925, linetype = 'twodash', color = "yellow") +
  #geom_text(label = "92.5%", x = 3.0, y = 0.96, color = "yellow", size = 6) +
  theme.type +
  theme(plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
ECDF_model_occurrence_ADAM_dark
```

# SFEI Data (uncorrected for blanks, do not use)
```{r eval=FALSE, include=FALSE}
# SFEI data
sfei <- read.csv("Concentration data/Datasets/SFEI.csv", na.strings = "N.A.") %>% 
  #filter(Below_LOD == "N") %>% 
  # align data
  mutate(x1M = 20,
         x2M = 5000) %>% 
  mutate(CF = CFfnx(x1M = x1M, x2M = x2M)) %>%  #create new column with correction factor 
  mutate(particle.L.master = CF * Particles.L_uncorrected)

# read in concentration data
samplessfei <- sfei %>% 
  mutate(Conc = particle.L.master)

#make new dataframe to plot both histograms together
sampleSimplesfei <- samplessfei %>%
  dplyr::select(Conc, Sample.Type) %>% 
  droplevels()

#make new dataframe to plot both histograms together
dfsfei <- rbind(sampleSimplesfei)#,food.dilution.simple)
```


## Exceedances
```{r eval=FALSE, include=FALSE}
#calculate exceedance
dfsfei_exceedance <- dfsfei %>% 
  mutate(aboveThreshold = factor(case_when(
    Conc < Threshold1 ~ "below Threshold 1",
    Conc >= Threshold1 & Conc < Threshold2 ~ "above Threshold 1",
    Conc >= Threshold2 & Conc < Threshold3 ~ "above Threshold 2",
    Conc >= Threshold3 & Conc < Threshold4 ~ "above Threshold 3",
    Conc >= Threshold4 ~ "above Threshold 4"
  )))

#give summary stat for exceedance
exceedance <- dfsfei_exceedance  %>%
  filter(Sample.Type == "sample") %>% 
  dplyr::select(c(Conc, aboveThreshold)) %>%
  group_by(aboveThreshold) %>%
  dplyr::summarize(n = n()) %>% 
  mutate(rel.freq = paste0(round(100 * n/sum(n), 0), "%"))

kable(exceedance)
```
## Histogram
```{r eval=FALSE, include=FALSE}
#generate plot
dfsfei_exceedance %>% 
  filter(Conc > 0) %>% 
  filter(Sample.Type == "sample") %>% 
  ggplot(aes(x = Conc, fill = aboveThreshold))+
  geom_histogram(aes(y = ..count../sum(..count..)),bins =30, alpha = 0.9, position = "identity") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.045), label = paste(Threshold1,"particles/L"),  color = "goldenrod2") +
  #geom_text(aes(x = Threshold1- 0.5*Threshold1, y = 0.055), label = ("Threshold1"),  color = "goldenrod2") +
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'goldenrod2', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'tomato1', size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'mediumvioletred', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'gray33', size = 1.3) +
  #geom_text(label = "Threshold 1", color = 'goldenrod2', x = log10(Threshold1) - 0.8*log10(Threshold1), y = 0.055, size = 6)+
#   geom_text(label = "Threshold 2", color = 'tomato1', x = log10(Threshold2) - 0.8*log10(Threshold2), y = 0.055, size = 6)+
 #  geom_text(label = "Threshold 3", color = 'mediumvioletred', x = log10(Threshold3) - 0.8*log10(Threshold3), y = 0.055, size = 6)+
  # geom_text(label = "Threshold 4", color = 'gray33', x = Threshold4 - 0.8*log10(Threshold4), y = 0.055, size = 6)+
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif, trans = "log10")+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  #coord_cartesian(xlim = c(0,100000000)) +
  #scale_x_continuous(labels = scales::scientific) +
   xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)")) +
  scale_y_continuous(name = "Relative Density", labels = scales::percent)+
  scale_fill_manual(values = c("below Threshold 1"= "cyan3", 
                               "above Threshold 1" = "goldenrod2", 
                               "above Threshold 2" = "tomato1",
                               "above Threshold 3" = "violetred4", 
                               "above Threshold 4" = "gray33")) +
  labs(title = "SF Bay and Sanctuary Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from sfeis et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))+
  theme.type +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```
## ECDF

```{r eval=FALSE, include=FALSE}
#ECDF by System
sfei %>% 
filter(System != "") %>% 
  ggplot(aes(x = particle.L.master)) + #, color = System))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.7) +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.3, size = 1.5, color = 'azure3') +
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = 'green', size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = 'yellow', size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'orange', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	'red', size = 1.3) +
  geom_text(label = "Threshold 1", color = 'green', x = Threshold1, y = 0.15, size = 6)+
  geom_text(label = "Threshold 2", color = 'yellow', x = Threshold2, y = 0.22, size = 6)+
  geom_text(label = "Threshold 3", color = 'orange', x = Threshold3, y = 0.35, size = 6)+
  geom_text(label = "Threshold 4", color = 'red', x = Threshold4, y = 0.45, size = 6)+
  ylab("Cumulative Density") +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
  scale_y_continuous(labels = scales::percent)+
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  theme.type +
  labs(title = "Global Surface Water Marine Microplastics Concentrations",
       subtitle = paste("Particles/L corrected to:",x1D_set, "to", x2D_set, "um"),
       caption = paste("Concentration data from sfeis et al (2019): 23 studies, 57 sampling locations, n = 377; corrected for size ranges. Alpha = ",alpha))+
  theme(
    plot.title = element_text(hjust = 0.5, size = 18),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

# Compare to all tox data
## MARINE
```{r ERM}
###define sizes for filtering and alignment##
# smaller size bin
small_tier_lower_size <- 1 #um
small_tier_upper_size <- 5000 #um #size to align to
upper.tissue.trans.size.um <- 83 #10 #um #set size for filtering data and x2M
# larger size bin
large_tier_lower_size <- 1 #um
large_tier_upper_size <- 5000 #um

#read in data
aoc_z <- readRDS(file = "Tox Data/aoc_z.Rda")

## First filter data with global filters
aoc_intermediate <- aoc_z %>% 
  filter(!environment %in% c("Terrestrial", "Not Reported"),
         org_f != "Bacterium",
         org_f != "Plant",
         effect.metric != "HONEC",
         tier_zero_tech_f == "Red Criteria Passed",
         tier_zero_risk_f == "Red Criteria Passed", #All thresholds must pass technical and risk red criteria
         risk.13 != 0 #Drop studies that received a score of 0 for endpoints criteria (this also drops studies that have not yet been scored) - KEEP THIS AFTER THE RED CRITERIA FILTERS  
         ) %>%  
  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) #makes it less confusing below
#### Ecologically Relevant Metric calculations ####


###function to derive correction factor (CF) from Koelmans et al (equation 2)
CFfnx = function(a, #default alpha from Koelmans et al (2020)
                 x2D, #set detault values to convert ranges to (1-5,000 um) #5mm is upper defuault 
                 x1D, #1 um is lower default size
                 x2M, x1M){
  CF = (x2D^(1-a)-x1D^(1-a))/(x2M^(1-a)-x1M^(1-a)) 
  return(CF)}

#### equations for mu_x_poly (note that there are three depending on certain alphas for limits of equation)
###### if alpha does not equal 2 #####
mux.polyfnx = function(a.x, 
                       x_UL, 
                       x_LL){
  mux.poly = ((1-a.x)/(2-a.x)) * ((x_UL^(2-a.x) - x_LL^(2-a.x))/(x_UL^(1-a.x) - x_LL^(1-a.x)))
  return(mux.poly)}

##### If alpha does equal 2 #####
mux.polyfnx.2 = function(x_UL,x_LL){
  mux.poly = (log(x_UL/x_LL))/(x_LL^(-1) - x_UL^-1)
  return(mux.poly)}

### Calculating max ingestible parameters ###
## function to calcualte min and max ingestible surface area ##
SAfnx = function(a, # a = 0.5 * length
                 b, # b = 0.5 * width
                 c # c = 0.5 * height (note that hieght is 0.67 * width)
){
  SA = 4*pi*(((a*b)^1.6 + (a*c)^1.6 + (b*c)^1.6) / 3)^(1/1.6)
  return(SA)}

## max ingestible volume ##
volumefnx = function(R, L){
  volume = 0.111667 * pi * R^2 * L^3 #assumes height = 0.67 * Width, and Width:Length ratio is 'R' (compartment-specific)
  return(volume)}

#max ingestible mass
massfnx = function(R, L, p){
  mass = p * #density (g/cm^3)
    0.111667 * pi * R^2 * L^3 * # volume (um^3): assumes height = 0.67 * Width, and Width:Length ratio is 'R' (compartment-specific)
    1/1e12 * 1e6 #correction factors
  return(mass)}

#max ingestible specific surface area
SSAfnx = function(sa, #surface area, calcaulted elsewhere
                  m){ #mass, calculated elsewhere
  SSA = sa/m
    return(SSA)}

## parametrization ##
# Define params for correction #
alpha = 2.07 #table s4 for marine surface water. length
# define parameters for power law coefficients
a.sa = 1.5 #marine surface area power law
a.v = 1.48 #a_V for marine surface water volume
a.m = 1.32 # upper limit fora_m for mass for marine surface water in table S4 
a.ssa = 1.98 # A_SSA for marine surface water

#define additional parameters for calculations based on averages in the environment
R.ave = 0.77 #average width to length ratio for microplastics in marine enviornment
p.ave = 1.10 #average density in marine surface water


##### SMALL TIER DATA ####
### Set size limits for tier
x2D_set = small_tier_upper_size #upper size range (default)
x1D_set = small_tier_lower_size #lower size range (default)
x1M_set = small_tier_lower_size #lower size range for measured

# calculate ERM for each species
aoc_final_small <- aoc_intermediate  %>% 
   # define upper size length for Translocation
  mutate(x2M = case_when(is.na(max.size.ingest.um) ~ upper.tissue.trans.size.um, 
                         max.size.ingest.um  < upper.tissue.trans.size.um ~  max.size.ingest.um,
                         max.size.ingest.um  > upper.tissue.trans.size.um ~ upper.tissue.trans.size.um)) %>%  #set to 10um for upper limit or max size ingest, whichever is smaller
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M, #upper ingestible size limit (width of particle)
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL = (EC_mono_p.particles.mL * mu.p.mono)/mu.p.poly) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M, #upper ingestible
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL = EC_poly_p.particles.mL * CF_bio) %>%  #aligned particle effect concentraiton (1-5000 um)
  
  #### Surface area ERM ####
  mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignment to ERM
  #calculate lower ingestible surface area
  mutate(x_LL_sa = SAfnx(a = 0.5 * x1D_set, 
                         b = 0.5 * R.ave * x1D_set, 
                         c = 0.5 * R.ave * 0.67 * x1D_set)) %>%  
  #calculate upper ingestible surface area
  mutate(x_UL_sa = SAfnx(a = 0.5 * x2M, 
                         b = 0.5 * R.ave *x2M, 
                         c = 0.5 * R.ave * 0.67 * x2M)) %>%  
  #calculate mu_x_poly for surface area
  mutate(mu.sa.poly = if(a.sa == 2){mux.polyfnx.2(x_UL_sa, x_LL_sa)} else if (a.sa != 2){mux.polyfnx(a.sa, x_UL_sa, x_LL_sa)}) %>% 
  #calculate polydisperse effect concentration for surface area (particles/mL)
  mutate(EC_poly_sa.particles.mL = (EC_mono_p.particles.mL * mu.sa.mono)/mu.sa.poly) %>%  
  #calculate environmentally realistic effect threshold
  mutate(EC_env_sa.particles.mL = EC_poly_sa.particles.mL * CF_bio) %>% 
  
  #### volume ERM ####
#define mu_x_mono for alignment to ERM
  mutate(mu.v.mono = particle.volume.um3) %>% 
  #calculate lower ingestible volume 
  mutate(x_LL_v = volumefnx(R = R.ave,
                            L = x1D_set)) %>%
  #calculate maximum ingestible volume 
  mutate(x_UL_v = volumefnx(R = R.ave,
                            L = x2M)) %>%  
  # calculate mu.v.poly
  mutate(mu.v.poly = if(a.v == 2){mux.polyfnx.2(x_UL_v, x_LL_v)} else if (a.v != 2){mux.polyfnx(a.v, x_UL_v, x_LL_v)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_v.particles.mL = (EC_mono_p.particles.mL * mu.v.mono)/mu.v.poly) %>%  
    #calculate environmentally realistic effect threshold
  mutate(EC_env_v.particles.mL = EC_poly_v.particles.mL * CF_bio) %>% 
  
  #### mass ERM ###
  #define mu_x_mono for alignment to ERM (ug)
  mutate(mu.m.mono = mass.per.particle.mg * 1000) %>% 
  #calculate lower ingestible mass
  mutate(x_LL_m = massfnx(R = R.ave, L = x1D_set, p = p.ave)) %>%  
  #calculate upper ingestible mass
  mutate(x_UL_m = massfnx(R = R.ave, L = x2M, p = p.ave)) %>%  
  # calculate mu.m.poly
  mutate(mu.m.poly = if(a.m == 2){mux.polyfnx.2(x_UL_m, x_LL_m)} else if (a.m != 2){mux.polyfnx(a.m, x_UL_m, x_LL_m)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_m.particles.mL = (EC_env_p.particles.mL * mu.m.mono)/mu.m.poly) %>%
  #calculate environmentally realistic effect threshold
  mutate(EC_env_m.particles.mL = EC_poly_m.particles.mL * CF_bio) %>% 
  
  ##### specific surface area ERM ####
  mutate(mu.ssa.mono = mu.sa.mono/mu.m.mono) %>% #define mu_x_mono for alignment to ERM (um^2/ug)
  #calculate lower ingestible SSA
  mutate(x_LL_ssa = SSAfnx(sa = x_LL_sa, #surface area
                           m = x_LL_m) #mass
         ) %>% 
  #calculate upper ingestible SSA  (um^2/ug)
  mutate(x_UL_ssa = SSAfnx(sa = x_UL_sa, #surface area
                           m = x_UL_m) #mass
         ) %>% 
  #calculate mu_x_poly for specific surface area
  mutate(mu.ssa.poly = if(a.ssa == 2){mux.polyfnx.2(x_UL_ssa, x_LL_ssa)} else if (a.ssa != 2){mux.polyfnx(a.ssa, x_UL_ssa, x_LL_ssa)}) %>% 
  #calculate polydisperse effect concentration for specific surface area (particles/mL)
  mutate(EC_poly_ssa.particles.mL = (EC_env_p.particles.mL * mu.ssa.mono)/mu.ssa.poly) %>% 
  #calculate environmentally realistic effect threshold
  mutate(EC_env_ssa.particles.mL = EC_poly_ssa.particles.mL * CF_bio)

##### LARGE TIER DATA ####
### Set size limits for large tier
x2D_set <- large_tier_upper_size #upper size range (default)
x1D_set <- large_tier_lower_size #lower size range (default)
x1M_set <- large_tier_lower_size #lower size range for measured
# calculate ERM for each species

aoc_final_large <- aoc_intermediate  %>% 
  # define upper size WIDTH for ingestion (based on average width:length ratio)
  mutate(x2M = case_when(is.na(max.size.ingest.um) ~ (1/R.ave) * x2D_set, #all calculations below occur for length. Width is R.ave * length, so correcting here makes width the max size ingest below
                         (max.size.ingest.um * (1/R.ave)) < x2D_set ~ ((1/R.ave) * max.size.ingest.um),
                         (max.size.ingest.um * (1/R.ave)) > x2D_set ~ (x2D_set * (1/R.ave))
                         )) %>%  #set to 10um for upper limit or max size ingest, whichever is smaller
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M, #upper ingestible size limit
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL = (EC_mono_p.particles.mL * mu.p.mono)/mu.p.poly) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M, #upper ingestible
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL = EC_poly_p.particles.mL * CF_bio) %>%  #aligned particle effect concentraiton (1-5000 um)
  #### Surface area ERM ####
  mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignment to ERM
  #calculate lower ingestible surface area
  mutate(x_LL_sa = SAfnx(a = 0.5 * x1D_set, 
                         b = 0.5 * R.ave * x1D_set, 
                         c = 0.5 * R.ave * 0.67 * x1D_set)) %>%  
  #calculate upper ingestible surface area
  mutate(x_UL_sa = SAfnx(a = 0.5 * x2M, 
                         b = 0.5 * R.ave *x2M, 
                         c = 0.5 * R.ave * 0.67 * x2M)) %>%  
  #calculate mu_x_poly for surface area
  mutate(mu.sa.poly = if(a.sa == 2){mux.polyfnx.2(x_UL_sa, x_LL_sa)} else if (a.sa != 2){mux.polyfnx(a.sa, x_UL_sa, x_LL_sa)}) %>% 
  #calculate polydisperse effect concentration for surface area (particles/mL)
  mutate(EC_poly_sa.particles.mL = (EC_mono_p.particles.mL * mu.sa.mono)/mu.sa.poly) %>%  
  #calculate environmentally realistic effect threshold
  mutate(EC_env_sa.particles.mL = EC_poly_sa.particles.mL * CF_bio) %>% 
  
  #### volume ERM ####
#define mu_x_mono for alignment to ERM
  mutate(mu.v.mono = particle.volume.um3) %>% 
  #calculate lower ingestible volume 
  mutate(x_LL_v = volumefnx(R = R.ave,
                            L = x1D_set)) %>%
  #calculate maximum ingestible volume 
  mutate(x_UL_v = volumefnx(R = R.ave,
                            L = x2M)) %>%  
  # calculate mu.v.poly
  mutate(mu.v.poly = if(a.v == 2){mux.polyfnx.2(x_UL_v, x_LL_v)} else if (a.v != 2){mux.polyfnx(a.v, x_UL_v, x_LL_v)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_v.particles.mL = (EC_mono_p.particles.mL * mu.v.mono)/mu.v.poly) %>%  
    #calculate environmentally realistic effect threshold
  mutate(EC_env_v.particles.mL = EC_poly_v.particles.mL * CF_bio) %>% 
  
  #### mass ERM ###
  #define mu_x_mono for alignment to ERM (ug)
  mutate(mu.m.mono = mass.per.particle.mg * 1000) %>% 
  #calculate lower ingestible mass
  mutate(x_LL_m = massfnx(R = R.ave, L = x1D_set, p = p.ave)) %>%  
  #calculate upper ingestible mass
  mutate(x_UL_m = massfnx(R = R.ave, L = x2M, p = p.ave)) %>%  
  # calculate mu.m.poly
  mutate(mu.m.poly = if(a.m == 2){mux.polyfnx.2(x_UL_m, x_LL_m)} else if (a.m != 2){mux.polyfnx(a.m, x_UL_m, x_LL_m)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_m.particles.mL = (EC_env_p.particles.mL * mu.m.mono)/mu.m.poly) %>%
  #calculate environmentally realistic effect threshold
  mutate(EC_env_m.particles.mL = EC_poly_m.particles.mL * CF_bio) %>% 
  
  ##### specific surface area ERM ####
  mutate(mu.ssa.mono = mu.sa.mono/mu.m.mono) %>% #define mu_x_mono for alignment to ERM (um^2/ug)
  #calculate lower ingestible SSA
  mutate(x_LL_ssa = SSAfnx(sa = x_LL_sa, #surface area
                           m = x_LL_m) #mass
         ) %>% 
  #calculate upper ingestible SSA  (um^2/ug)
  mutate(x_UL_ssa = SSAfnx(sa = x_UL_sa, #surface area
                           m = x_UL_m) #mass
         ) %>% 
  #calculate mu_x_poly for specific surface area
  mutate(mu.ssa.poly = if(a.ssa == 2){mux.polyfnx.2(x_UL_ssa, x_LL_ssa)} else if (a.ssa != 2){mux.polyfnx(a.ssa, x_UL_ssa, x_LL_ssa)}) %>% 
  #calculate polydisperse effect concentration for specific surface area (particles/mL)
  mutate(EC_poly_ssa.particles.mL = (EC_env_p.particles.mL * mu.ssa.mono)/mu.ssa.poly) %>% 
  #calculate environmentally realistic effect threshold
  mutate(EC_env_ssa.particles.mL = EC_poly_ssa.particles.mL * CF_bio)
```

### Plot Aligned Data
Combine tox dataset with occurence data
```{r}
volume <- aoc_final_large %>% 
  # tier 3/4 filters
  filter(risk.13 != 1,
         bio_f %in% c("Organism", "Population")) %>% 
  filter(Group != "Algae") %>% 
  filter(environment == "Marine") %>% 
         mutate(Conc = EC_env_v.particles.mL / (af.time * af.noec)) %>%  
         drop_na(Conc) %>% 
         mutate(Conc = Conc * 1000) %>% 
  filter(between(size.length.um.used.for.conversions, 1, 5000))  %>% 
  mutate(tox.occurrence = "tox") %>% 
  dplyr::select(Conc, tox.occurrence)

global <- globalData %>% 
  filter(System %in% c("Ocean", "Estuary")) %>% 
  mutate(tox.occurrence = "occurence") %>% 
  dplyr::select(Conc, tox.occurrence)

tox.occurence <- rbind.data.frame(volume, global)

#build plot
aligned.tox.occurence.plot <- tox.occurence %>% 
  ggplot(aes(x = Conc, fill = tox.occurrence)) +
  #geom_histogram(bins = 40) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", labels = c("Occurence", "NOECs")) +
  scale_x_log10(name = paste("Particles/L (corrected to:",x1D_set, "to", x2D_set, "um)"),
                             breaks = scales::log_breaks(),
                             labels = comma_signif) +
  ylab("Normalized Abundance") +
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  #labs(#title = "Global Surface Water Marine Microplastics Concentrations vs. Tox Data",
       #subtitle = "Tox data filtered for quality & risk: organism, Population, fitness endpoints",
     #  caption = paste("(26 occurence studies, 37 tox studies, Alpha = ",alpha,")"))+
  #dark_theme_minimal(base_size = 25) +
 theme_bw(base_size = 15) +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        #axis.text =  element_text(size = 16),
        #legend.text = element_text(size =18),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 14))

aligned.tox.occurence.plot
```
### Plot Un-aligned Data
Combine tox dataset with occurence data
```{r}
unaligned.tox <- aoc_final_large %>% 
  # tier 3/4 filters
  filter(risk.13 != 1,
         bio_f %in% c("Organism", "Population")) %>% 
         mutate(Conc = dose.particles.mL.master / (af.time * af.noec)) %>%  
         drop_na(Conc) %>% 
         mutate(Conc = Conc * 1000) %>% 
  filter(Group != "Algae") %>% 
  filter(environment == "Marine") %>% 
  filter(between(size.length.um.used.for.conversions, 1, 5000))  %>% 
  mutate(tox.occurrence = "tox") %>% 
  dplyr::select(Conc, tox.occurrence)

# revert aligned occurence data to unaligned
unaligned.global <- globalData %>% 
  filter(System %in% c("Ocean", "Estuary")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  dplyr::select(unaligned) %>% 
  rename(Conc = unaligned) %>% 
  mutate(tox.occurrence = "occurence")


unaligned.tox.occurence <- rbind.data.frame(unaligned.tox, unaligned.global)


####define overlap ###
#bin the data using density
tox.unaligned <- unaligned.tox.occurence %>% 
  filter(tox.occurrence == "tox") %>% 
    dplyr::select(Conc)

occurrence.unaligned <- unaligned.tox.occurence %>% 
  filter(tox.occurrence == "occurence") %>% 
    dplyr::select(Conc)


# 
# ## METHOD 1
# tox.unaligned.density <- density(tox.unaligned$Conc)
# occurrence.unaligned.density <- density(occurrence.unaligned$Conc)
# 
# idx <- (occurrence.unaligned.density$y < tox.unaligned.density$y) &
# (occurrence.unaligned.density$x > min(occurrence.unaligned.density$x)) &
# (occurrence.unaligned.density$x < max(occurrence.unaligned.density$x))
# #overlap
# unaligned_intersection <- min(occurrence.unaligned.density$x[idx])  
# unaligned_intersection
# 
# ##Method 2
# #remotes::install_github("andrewheiss/reconPlots")
# require(reconPlots)
# occurrence.unaligned.df <- data.frame(x = occurrence.unaligned.density$x, 
#                                       y =occurrence.unaligned.density$y)
# 
# tox.unaligned.df <- data.frame(x = tox.unaligned.density$x, 
#                                       y =tox.unaligned.density$y)
# 
# curve_intersect(occurrence.unaligned.df, tox.unaligned.df, empirical = TRUE)
```

```{r}
##### Genereate PLot###
unaligned.tox.occurence.plot <- unaligned.tox.occurence %>% 
  ggplot(aes(x = Conc, fill = tox.occurrence)) +
  #geom_histogram(bins = 40) +
  geom_density(alpha = 0.6) +
  #geom_vline(xintercept = unaligned_intersection, linetype = 'dashed', size = 0.5, color = "black") +
  #annotate(geom = "text", label = round(unaligned_intersection, 2), color = "black", 
   #        x = unaligned_intersection - 1, y = 0.1, size = 4, angle = 90) +
  scale_fill_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", labels = c("Occurence", "NOECs")) +
  scale_x_log10(name = paste("Particles/L (unaligned)"),
                             breaks = scales::log_breaks(),
                             labels = comma_signif) +
  ylab("Normalized Abundance") +
  # annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  # labs(#title = "Global Surface Water Marine Microplastics Concentrations vs. Tox Data",
  #      #subtitle = "Tox data filtered for quality & risk: organism, Population, fitness endpoints",
  #      caption = paste("(24 occurence studies, 37 tox studies, Alpha = ",alpha,")"))+
  #dark_theme_minimal(base_size = 25) +
  theme_bw(base_size = 15) +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        #axis.text =  element_text(size = 16),
        #legend.text = element_text(size =18),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 14))

unaligned.tox.occurence.plot
```
#### Manuscript Figure
```{r}
overlap_figures <- ggarrange(unaligned.tox.occurence.plot, aligned.tox.occurence.plot,
                        labels = c("A", "B"), common.legend = TRUE,
                        legend = "bottom",
                        ncol = 2, nrow = 1)

plot(overlap_figures)

ggsave(plot =overlap_figures,
       filename = "overlap_figure_SI-Fig4.tiff",
       path = "./Concentration data/Threshold_Manuscript_Figs/", 
       width = 20, height = 12, units = "in",
       dpi = 300)
```
### ECDF
```{r}
vals <- c("Unaligned" = "black", "Aligned" = "blue")

ECDF_marine <- globalData %>% 
   filter(System %in% c("Ocean", "Estuary")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  ggplot(aes(x = unaligned)) + 
  #unaligned
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #aligned
  stat_ecdf(aes(x = Conc, color = "Aligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = Conc, color = "Aligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
 
  #thresholds
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "1", color = "#00798c" , x = log10(Threshold1) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "2", color = "#66a182" , x = log10(Threshold2) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "3", color = '#edae49', x = log10(Threshold3) + 0.2, y = 0.1, size = 5)+
  geom_text(label = "4", color = "#d1495b", x = log10(Threshold4) + 0.2, y = 0.1, size = 5)+
  #aesthetics
   ylab("Cumulative Density") +
  xlab(paste0("Particles/L"))+
  scale_y_continuous(labels = scales::percent) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif) +
   #color scale
  scale_color_manual(name = "alignment",
                     values = vals) +
  #coord_trans(x = "log10") +
  
  #annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  theme.type +
  theme(legend.title = element_blank())

ECDF_marine
```



## Table comparing overlap
```{r}
#require(overlapping)

#can't install package

# #bin the data using density
# tox.unaligned <- unaligned.tox.occurence %>% 
#   filter(tox.occurrence == "tox") %>% 
#     dplyr::select(Conc)
# 
# occurrence.unaligned <- unaligned.tox.occurence %>% 
#   filter(tox.occurrence == "occurence") %>% 
#     dplyr::select(Conc)
# # create list with different distributions
# unaligned <- list(tox = tox.unaligned$Conc, occurrence = occurrence.unaligned$Conc)
# out <- overlap(unaligned, plot = TRUE)
# out$OV #estimate overlapped areas
```

```{r}
require(ggplot2)
require(reshape2)

# Melt and plot
munaligned.tox.occurence <- melt(unaligned.tox.occurence)

ggplot(munaligned.tox.occurence) +
  geom_density(aes(x = log10(value), color = tox.occurrence))

# Find points that intersect  
#intersect(df$x, df$y)
```


# Tox Sizes vs. Monitoring Cut-offs
## Histograms
```{r}
unaligned.tox.2 <- aoc_final_large %>% 
  # tier 3/4 filters
  #filter(risk.13 != 1,
         #bio_f %in% c("Organism", "Population")) %>% 
         mutate(Conc = dose.particles.mL.master / (af.time * af.noec)) %>%  
         drop_na(Conc) %>% 
         mutate(Conc = Conc * 1000) %>% 
 # filter(between(size.length.um.used.for.conversions, 1, 5000))  %>% 
  mutate(tox.occurrence = "tox") %>% 
  mutate(x1M = size.length.um.used.for.conversions) %>% 
  dplyr::select(x1M, tox.occurrence)

# extract size cutoff from monitoring data
unaligned.global.2 <- dfADAM %>% 
  #mutate(unaligned = particle.L.master / CF) %>%
  dplyr::select(x1M) %>% 
  #rename(Conc = unaligned) %>% 
  mutate(tox.occurrence = "occurence")
  
#join datasets for plotting
sizes.tox.occurence <- rbind.data.frame(unaligned.tox.2, unaligned.global.2)

#### PLOTTING ####
sizes.tox.occurence %>% 
   ggplot(aes(x = x1M, fill = tox.occurrence)) +
  #geom_histogram(bins = 40) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", 
                    labels = c("Lower size limits for microplastics sampling",
                               "Sizes of Microplastics tested in ecotoxicology")) +
  scale_x_log10(name =  bquote("size ("~mu*"m)"),
                             breaks = scales::log_breaks(),
                             labels = comma_signif) +
  ylab("Relative Frequency") +
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  # labs(title = "Microplastics Monitoring Lower Size Limits vs. Sizes of Microplastics Tested in Ecotoxicology Studies"#,
  #      #caption = paste("(24 occurence studies, 37 tox studies, Alpha = ",alpha,")")
  #      ) +
  dark_theme_minimal() +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
        #axis.text =  element_text(size = 16),
        legend.text = element_text(size =12),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "vertical",
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 12))
```

## ECDFs

```{r}

# get lower limit for occurence data for plotting
lower.size <- sizes.tox.occurence %>% 
  filter(tox.occurrence == "occurence") %>% 
  summarize("lower" = min(x1M))

#build ECDF
sizes.tox.occurence %>% 
ggplot(aes(x = x1M, color = tox.occurrence))+
  stat_ecdf(geom = "point", size = 1, alpha = 0.9, pad = FALSE) +
  stat_ecdf(geom = "step", linetype = 'solid', alpha = 0.7, size = 1.5, pad = FALSE) +
  scale_color_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", 
                    labels = c("Lower size limits for microplastics sampling",
                               "Sizes of Microplastics tested in ecotoxicology")) +
  ylab("Cumulative Frequency") +
  xlab(bquote("size ("~mu*"m)")) +
  # lower size dotted lines
  geom_segment(x = lower.size[[1]], xend = lower.size[[1]], y = 0.13, yend = 0, 
  linetype = 'dashed', color = '#00798c', size = 1) +
  geom_segment(x = 0.001, xend = lower.size[[1]], y = 0.13, yend = 0.13, 
  linetype = 'dashed', color = '#00798c', size = 1) +
  scale_y_continuous(labels = scales::percent) +
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  labs(title = "Microplastics Monitoring Lower Size Limits vs. Sizes of Microplastics Tested in Ecotoxicology Studies") +
  dark_theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 16),
        axis.title = element_text(size = 16),
    axis.text = element_text(size = 14),
        #axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
    legend.position = "top",
    legend.justification = "left",
    legend.direction = "vertical",
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 12))
```

# Are sampling sizes and concentrations related?
The following analysis tests whether the alignments are accurate.
```{r}
oceanData <- globalData %>% 
  filter(System == "Ocean")


globalData %>% 
  filter(System == "Ocean") %>% 
  ggplot(aes(x = x1M, y = Conc, color = Sampling.location)) +
  geom_point() +
  scale_x_log10(name = "Lower Size Limit of Detection") +
  scale_y_log10(name = "1 - 5,000 µm Aligned Concentration (Particles/L)") +
  dark_theme_bw()
  
```
```{r}
globalData %>% 
  filter(System == "Ocean") %>% 
  mutate(Sampling.location = fct_reorder(Sampling.location, Conc, .fun = median)) %>% 
  ggplot(aes(x = Conc, y = Sampling.location, color = Sampling.location)) +
   geom_vline(xintercept = Threshold1, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "Tier 1", color = "#00798c" , x = Threshold1, y = 14, size = 5)+
  geom_text(label = "Tier 2", color = "#66a182" , x = Threshold2, y = 13.5, size = 5)+
  geom_text(label = "Tier 3", color = '#edae49', x = Threshold3, y = 13, size = 5)+
  geom_text(label = "Tier 4", color = "#d1495b", x = Threshold4, y = 12.5, size = 5)+
  geom_boxplot() +
  geom_point() +
  xlab(paste0("Particles/L (" ,x1D_set, " to ", x2D_set, "um)"))+
  coord_trans(x = "log10") +
  scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif)+
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  theme.type +
  dark_theme_bw()
```


## FRESHWATER
```{r ERM}
###define sizes for filtering and alignment##
# smaller size bin
small_tier_lower_size <- 1 #um
small_tier_upper_size <- 5000 #um #size to align to
upper.tissue.trans.size.um <- 83 #10 #um #set size for filtering data and x2M
# larger size bin
large_tier_lower_size <- 1 #um
large_tier_upper_size <- 5000 #um

#read in data
aoc_z <- readRDS(file = "Tox Data/aoc_z.Rda")

## First filter data with global filters
aoc_intermediate <- aoc_z %>% 
  #filter(!environment %in% c("Terrestrial", "Not Reported"),
  filter(environment == "Freshwater",
         org_f != "Bacterium",
         org_f != "Plant",
         effect.metric != "HONEC",
         tier_zero_tech_f == "Red Criteria Passed",
         tier_zero_risk_f == "Red Criteria Passed", #All thresholds must pass technical and risk red criteria
         risk.13 != 0 #Drop studies that received a score of 0 for endpoints criteria (this also drops studies that have not yet been scored) - KEEP THIS AFTER THE RED CRITERIA FILTERS  
         ) %>%  
  mutate(max.size.ingest.um = 1000 * max.size.ingest.mm) #makes it less confusing below
#### Ecologically Relevant Metric calculations ####
## parametrization ##
# Define params for correction #

alpha = 2.64 #table s4 for freshwater surface water. length
# define parameters for power law coefficients
a.sa = 2.00 #freshwater surface area power law
a.v = 1.68 #a_V for Freshwater surface water volume
a.m = 1.65 # upper limit fora_m for mass for freshwater surface water in table S4 
a.ssa = 2.71 # A_SSA for freshwater surface water

#define additional parameters for calculations based on averages in the environment
R.ave = 0.67 #average width to length ratio for microplastics in surface freshwater
p.ave = 1.04 #average density in freshwater surface water

##### SMALL TIER DATA ####
### Set size limits for tier
x2D_set = small_tier_upper_size #upper size range (default)
x1D_set = small_tier_lower_size #lower size range (default)
x1M_set = small_tier_lower_size #lower size range for measured

# calculate ERM for each species
aoc_final_small <- aoc_intermediate  %>% 
   # define upper size length for Translocation
  mutate(x2M = case_when(is.na(max.size.ingest.um) ~ upper.tissue.trans.size.um, 
                         max.size.ingest.um  < upper.tissue.trans.size.um ~  max.size.ingest.um,
                         max.size.ingest.um  > upper.tissue.trans.size.um ~ upper.tissue.trans.size.um)) %>%  #set to 10um for upper limit or max size ingest, whichever is smaller
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M, #upper ingestible size limit (width of particle)
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL = (EC_mono_p.particles.mL * mu.p.mono)/mu.p.poly) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M, #upper ingestible
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL = EC_poly_p.particles.mL * CF_bio) %>%  #aligned particle effect concentraiton (1-5000 um)
  
  #### Surface area ERM ####
  mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignment to ERM
  #calculate lower ingestible surface area
  mutate(x_LL_sa = SAfnx(a = 0.5 * x1D_set, 
                         b = 0.5 * R.ave * x1D_set, 
                         c = 0.5 * R.ave * 0.67 * x1D_set)) %>%  
  #calculate upper ingestible surface area
  mutate(x_UL_sa = SAfnx(a = 0.5 * x2M, 
                         b = 0.5 * R.ave *x2M, 
                         c = 0.5 * R.ave * 0.67 * x2M)) %>%  
  #calculate mu_x_poly for surface area
  mutate(mu.sa.poly = if(a.sa == 2){mux.polyfnx.2(x_UL_sa, x_LL_sa)} else if (a.sa != 2){mux.polyfnx(a.sa, x_UL_sa, x_LL_sa)}) %>% 
  #calculate polydisperse effect concentration for surface area (particles/mL)
  mutate(EC_poly_sa.particles.mL = (EC_mono_p.particles.mL * mu.sa.mono)/mu.sa.poly) %>%  
  #calculate environmentally realistic effect threshold
  mutate(EC_env_sa.particles.mL = EC_poly_sa.particles.mL * CF_bio) %>% 
  
  #### volume ERM ####
#define mu_x_mono for alignment to ERM
  mutate(mu.v.mono = particle.volume.um3) %>% 
  #calculate lower ingestible volume 
  mutate(x_LL_v = volumefnx(R = R.ave,
                            L = x1D_set)) %>%
  #calculate maximum ingestible volume 
  mutate(x_UL_v = volumefnx(R = R.ave,
                            L = x2M)) %>%  
  # calculate mu.v.poly
  mutate(mu.v.poly = if(a.v == 2){mux.polyfnx.2(x_UL_v, x_LL_v)} else if (a.v != 2){mux.polyfnx(a.v, x_UL_v, x_LL_v)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_v.particles.mL = (EC_mono_p.particles.mL * mu.v.mono)/mu.v.poly) %>%  
    #calculate environmentally realistic effect threshold
  mutate(EC_env_v.particles.mL = EC_poly_v.particles.mL * CF_bio) %>% 
  
  #### mass ERM ###
  #define mu_x_mono for alignment to ERM (ug)
  mutate(mu.m.mono = mass.per.particle.mg * 1000) %>% 
  #calculate lower ingestible mass
  mutate(x_LL_m = massfnx(R = R.ave, L = x1D_set, p = p.ave)) %>%  
  #calculate upper ingestible mass
  mutate(x_UL_m = massfnx(R = R.ave, L = x2M, p = p.ave)) %>%  
  # calculate mu.m.poly
  mutate(mu.m.poly = if(a.m == 2){mux.polyfnx.2(x_UL_m, x_LL_m)} else if (a.m != 2){mux.polyfnx(a.m, x_UL_m, x_LL_m)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_m.particles.mL = (EC_env_p.particles.mL * mu.m.mono)/mu.m.poly) %>%
  #calculate environmentally realistic effect threshold
  mutate(EC_env_m.particles.mL = EC_poly_m.particles.mL * CF_bio) %>% 
  
  ##### specific surface area ERM ####
  mutate(mu.ssa.mono = mu.sa.mono/mu.m.mono) %>% #define mu_x_mono for alignment to ERM (um^2/ug)
  #calculate lower ingestible SSA
  mutate(x_LL_ssa = SSAfnx(sa = x_LL_sa, #surface area
                           m = x_LL_m) #mass
         ) %>% 
  #calculate upper ingestible SSA  (um^2/ug)
  mutate(x_UL_ssa = SSAfnx(sa = x_UL_sa, #surface area
                           m = x_UL_m) #mass
         ) %>% 
  #calculate mu_x_poly for specific surface area
  mutate(mu.ssa.poly = if(a.ssa == 2){mux.polyfnx.2(x_UL_ssa, x_LL_ssa)} else if (a.ssa != 2){mux.polyfnx(a.ssa, x_UL_ssa, x_LL_ssa)}) %>% 
  #calculate polydisperse effect concentration for specific surface area (particles/mL)
  mutate(EC_poly_ssa.particles.mL = (EC_env_p.particles.mL * mu.ssa.mono)/mu.ssa.poly) %>% 
  #calculate environmentally realistic effect threshold
  mutate(EC_env_ssa.particles.mL = EC_poly_ssa.particles.mL * CF_bio)

##### LARGE TIER DATA ####
### Set size limits for large tier
x2D_set <- large_tier_upper_size #upper size range (default)
x1D_set <- large_tier_lower_size #lower size range (default)
x1M_set <- large_tier_lower_size #lower size range for measured
# calculate ERM for each species

aoc_final_large <- aoc_intermediate  %>% 
  # define upper size WIDTH for ingestion (based on average width:length ratio)
  mutate(x2M = case_when(is.na(max.size.ingest.um) ~ (1/R.ave) * x2D_set, #all calculations below occur for length. Width is R.ave * length, so correcting here makes width the max size ingest below
                         (max.size.ingest.um * (1/R.ave)) < x2D_set ~ ((1/R.ave) * max.size.ingest.um),
                         (max.size.ingest.um * (1/R.ave)) > x2D_set ~ (x2D_set * (1/R.ave))
                         )) %>%  #set to 10um for upper limit or max size ingest, whichever is smaller
 # calculate effect threshold for particles
  mutate(EC_mono_p.particles.mL = dose.particles.mL.master) %>% 
  mutate(mu.p.mono = 1) %>% #mu_x_mono is always 1 for particles to particles
  mutate(mu.p.poly = mux.polyfnx(a.x = alpha, #alpha for particles
                                 x_UL= x2M, #upper ingestible size limit
                                 x_LL = x1M_set)) %>% 
  # polydisperse effect threshold for particles
  mutate(EC_poly_p.particles.mL = (EC_mono_p.particles.mL * mu.p.mono)/mu.p.poly) %>% 
   #calculate CF_bio for all conversions
  mutate(CF_bio = CFfnx(x1M = x1M_set,#lower size bin
                        x2M = x2M, #upper ingestible
                        x1D = x1D_set, #default
                        x2D = x2D_set,  #default
                        a = alpha)) %>%  
  ## Calculate environmentally relevant effect threshold for particles
  mutate(EC_env_p.particles.mL = EC_poly_p.particles.mL * CF_bio) %>%  #aligned particle effect concentraiton (1-5000 um)
  #### Surface area ERM ####
  mutate(mu.sa.mono = particle.surface.area.um2) %>% #define mu_x_mono for alignment to ERM
  #calculate lower ingestible surface area
  mutate(x_LL_sa = SAfnx(a = 0.5 * x1D_set, 
                         b = 0.5 * R.ave * x1D_set, 
                         c = 0.5 * R.ave * 0.67 * x1D_set)) %>%  
  #calculate upper ingestible surface area
  mutate(x_UL_sa = SAfnx(a = 0.5 * x2M, 
                         b = 0.5 * R.ave *x2M, 
                         c = 0.5 * R.ave * 0.67 * x2M)) %>%  
  #calculate mu_x_poly for surface area
  mutate(mu.sa.poly = if(a.sa == 2){mux.polyfnx.2(x_UL_sa, x_LL_sa)} else if (a.sa != 2){mux.polyfnx(a.sa, x_UL_sa, x_LL_sa)}) %>% 
  #calculate polydisperse effect concentration for surface area (particles/mL)
  mutate(EC_poly_sa.particles.mL = (EC_mono_p.particles.mL * mu.sa.mono)/mu.sa.poly) %>%  
  #calculate environmentally realistic effect threshold
  mutate(EC_env_sa.particles.mL = EC_poly_sa.particles.mL * CF_bio) %>% 
  
  #### volume ERM ####
#define mu_x_mono for alignment to ERM
  mutate(mu.v.mono = particle.volume.um3) %>% 
  #calculate lower ingestible volume 
  mutate(x_LL_v = volumefnx(R = R.ave,
                            L = x1D_set)) %>%
  #calculate maximum ingestible volume 
  mutate(x_UL_v = volumefnx(R = R.ave,
                            L = x2M)) %>%  
  # calculate mu.v.poly
  mutate(mu.v.poly = if(a.v == 2){mux.polyfnx.2(x_UL_v, x_LL_v)} else if (a.v != 2){mux.polyfnx(a.v, x_UL_v, x_LL_v)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_v.particles.mL = (EC_mono_p.particles.mL * mu.v.mono)/mu.v.poly) %>%  
    #calculate environmentally realistic effect threshold
  mutate(EC_env_v.particles.mL = EC_poly_v.particles.mL * CF_bio) %>% 
  
  #### mass ERM ###
  #define mu_x_mono for alignment to ERM (ug)
  mutate(mu.m.mono = mass.per.particle.mg * 1000) %>% 
  #calculate lower ingestible mass
  mutate(x_LL_m = massfnx(R = R.ave, L = x1D_set, p = p.ave)) %>%  
  #calculate upper ingestible mass
  mutate(x_UL_m = massfnx(R = R.ave, L = x2M, p = p.ave)) %>%  
  # calculate mu.m.poly
  mutate(mu.m.poly = if(a.m == 2){mux.polyfnx.2(x_UL_m, x_LL_m)} else if (a.m != 2){mux.polyfnx(a.m, x_UL_m, x_LL_m)}) %>% 
  #calculate polydisperse effect concentration for volume (particles/mL)
  mutate(EC_poly_m.particles.mL = (EC_env_p.particles.mL * mu.m.mono)/mu.m.poly) %>%
  #calculate environmentally realistic effect threshold
  mutate(EC_env_m.particles.mL = EC_poly_m.particles.mL * CF_bio) %>% 
  
  ##### specific surface area ERM ####
  mutate(mu.ssa.mono = mu.sa.mono/mu.m.mono) %>% #define mu_x_mono for alignment to ERM (um^2/ug)
  #calculate lower ingestible SSA
  mutate(x_LL_ssa = SSAfnx(sa = x_LL_sa, #surface area
                           m = x_LL_m) #mass
         ) %>% 
  #calculate upper ingestible SSA  (um^2/ug)
  mutate(x_UL_ssa = SSAfnx(sa = x_UL_sa, #surface area
                           m = x_UL_m) #mass
         ) %>% 
  #calculate mu_x_poly for specific surface area
  mutate(mu.ssa.poly = if(a.ssa == 2){mux.polyfnx.2(x_UL_ssa, x_LL_ssa)} else if (a.ssa != 2){mux.polyfnx(a.ssa, x_UL_ssa, x_LL_ssa)}) %>% 
  #calculate polydisperse effect concentration for specific surface area (particles/mL)
  mutate(EC_poly_ssa.particles.mL = (EC_env_p.particles.mL * mu.ssa.mono)/mu.ssa.poly) %>% 
  #calculate environmentally realistic effect threshold
  mutate(EC_env_ssa.particles.mL = EC_poly_ssa.particles.mL * CF_bio)
```

### Plot Aligned Data
Combine tox dataset with occurence data
```{r}
volume <- aoc_final_large %>% 
  # tier 3/4 filters
  filter(risk.13 != 1,
         bio_f %in% c("Organism", "Population")) %>% 
  filter(Group != "Algae") %>% 
  filter(environment == "Freshwater") %>% 
         mutate(Conc = EC_env_v.particles.mL / (af.time * af.noec)) %>%  
         drop_na(Conc) %>% 
         mutate(Conc = Conc * 1000) %>% 
  filter(between(size.length.um.used.for.conversions, 1, 5000))  %>% 
  mutate(tox.occurrence = "tox") %>% 
  dplyr::select(Conc, tox.occurrence)

# occureence data
#make new dataframe to plot both histograms together
freshwater_aligned <- globalData %>%
  filter(sample.matrix.specific == "surface water") %>% 
  filter(System == c("Lake", "River")) %>% 
  drop_na(Conc) %>% 
  droplevels() %>% 
  mutate(tox.occurrence = "occurence") %>% 
  dplyr::select(Conc, tox.occurrence)

freshwater_aligned.tox.occurence <- rbind.data.frame(volume, freshwater_aligned)

#build plot
freshwater.aligned.tox.occurence.plot <- freshwater_aligned.tox.occurence %>% 
  ggplot(aes(x = Conc, fill = tox.occurrence)) +
  #geom_histogram(bins = 40) +
  geom_density(alpha = 0.6) +
  scale_fill_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", labels = c("Occurence", "NOECs")) +
  scale_x_log10(name = paste("Particles/L (corrected to:",x1D_set, "to", x2D_set, "um)"),
                             breaks = scales::log_breaks(),
                             labels = comma_signif) +
  ylab("Normalized Abundance") +
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  # labs(#title = "Global Surface Water Freshwater Microplastics Concentrations vs. Tox Data",
  #      #subtitle = "Tox data filtered for quality & risk: organism, Population, fitness endpoints",
  #      caption = paste("(70 occurence studies, 14 tox studies, Alpha = ",alpha,")"))+
  #dark_theme_minimal(base_size = 25) +
 theme_bw(base_size = 15) +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        #axis.text =  element_text(size = 16),
        #legend.text = element_text(size =18),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 14))

freshwater.aligned.tox.occurence.plot
```
### Plot Un-aligned Data
Combine tox dataset with occurence data
```{r}
unaligned.tox <- aoc_final_large %>% 
  # tier 3/4 filters
  filter(risk.13 != 1,
         bio_f %in% c("Organism", "Population")) %>% 
         mutate(Conc = dose.particles.mL.master / (af.time * af.noec)) %>%  
         drop_na(Conc) %>% 
         mutate(Conc = Conc * 1000) %>% 
  filter(Group != "Algae") %>% 
  filter(environment == "Freshwater") %>% 
  filter(between(size.length.um.used.for.conversions, 1, 5000))  %>% 
  mutate(tox.occurrence = "tox") %>% 
  dplyr::select(Conc, tox.occurrence)

# revert aligned occurence data to unaligned
unaligned.global <- globalData %>%
  filter(sample.matrix.specific == "surface water") %>% 
  filter(System == c("Lake", "River")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  dplyr::select(unaligned) %>% 
  rename(Conc = unaligned) %>% 
  mutate(tox.occurrence = "occurence")



unaligned.tox.occurence <- rbind.data.frame(unaligned.tox, unaligned.global)


####define overlap ###
#bin the data using density
tox.unaligned <- unaligned.tox.occurence %>% 
  filter(tox.occurrence == "tox") %>% 
    dplyr::select(Conc)

occurrence.unaligned <- unaligned.tox.occurence %>% 
  filter(tox.occurrence == "occurence") %>% 
    dplyr::select(Conc)



## METHOD 1
tox.unaligned.density <- density(tox.unaligned$Conc)
occurrence.unaligned.density <- density(occurrence.unaligned$Conc)

idx <- (occurrence.unaligned.density$y < tox.unaligned.density$y) &
(occurrence.unaligned.density$x > min(occurrence.unaligned.density$x)) &
(occurrence.unaligned.density$x < max(occurrence.unaligned.density$x))
#overlap
unaligned_intersection <- min(occurrence.unaligned.density$x[idx])  
unaligned_intersection

##Method 2
#remotes::install_github("andrewheiss/reconPlots")
require(reconPlots)
occurrence.unaligned.df <- data.frame(x = occurrence.unaligned.density$x, 
                                      y =occurrence.unaligned.density$y)

tox.unaligned.df <- data.frame(x = tox.unaligned.density$x, 
                                      y =tox.unaligned.density$y)

#curve_intersect(occurrence.unaligned.df, tox.unaligned.df, empirical = TRUE)
```

```{r}
##### Genereate PLot###
freshwater.unaligned.tox.occurence.plot <- unaligned.tox.occurence %>% 
  ggplot(aes(x = Conc, fill = tox.occurrence)) +
  #geom_histogram(bins = 40) +
  geom_density(alpha = 0.6) +
  #geom_vline(xintercept = unaligned_intersection, linetype = 'dashed', size = 0.5, color = "black") +
  #annotate(geom = "text", label = round(unaligned_intersection, 2), color = "black", 
   #        x = unaligned_intersection - 1, y = 0.1, size = 4, angle = 90) +
  scale_fill_manual(values = cal_palette(name = "superbloom3", type = "discrete"),
                    name = "Tox or Occurence Data", labels = c("Occurence", "NOECs")) +
  scale_x_log10(name = paste("Particles/L (unaligned)"),
                             breaks = scales::log_breaks(),
                             labels = comma_signif) +
  ylab("Normalized Abundance") +
  annotation_logticks(sides = "b")+ #log scale rick marks on bottom
  # labs(#title = "Global Surface Water Freshwater Microplastics Concentrations vs. Tox Data",
  #      #subtitle = "Tox data filtered for quality & risk: organism, Population, fitness endpoints",
  #      caption = paste("(70 occurence studies, 14 tox studies, Alpha = ",alpha,")"))+
  #dark_theme_minimal(base_size = 25) +
  theme_bw(base_size = 15) +
  theme(#legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        #axis.text =  element_text(size = 16),
        #legend.text = element_text(size =18),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14),
    plot.caption = element_text(hjust = 0.9, size = 14))

freshwater.unaligned.tox.occurence.plot
```
#### Manuscript Figure
```{r}
overlap_figures_marine_freshwater <- ggarrange(unaligned.tox.occurence.plot, aligned.tox.occurence.plot,
                                               freshwater.unaligned.tox.occurence.plot, freshwater.aligned.tox.occurence.plot,
                        labels = c("A", "B", "C", "D"), common.legend = TRUE,
                        legend = "bottom",
                        ncol = 2, nrow = 2)

plot(overlap_figures_marine_freshwater)

ggsave(plot =overlap_figures_marine_freshwater,
       filename = "overlap_figure_marine_freshwater_SI-Fig5.tiff",
       path = "./Concentration data/Threshold_Manuscript_Figs/", 
       width = 12, height = 8, units = "in",
       dpi = 300)
```
## ECDF
```{r}
vals <- c("Unaligned" = "black", "Aligned" = "blue")

ECDF_freshwater <- globalData %>% 
  filter(sample.matrix.specific == "surface water") %>% 
   filter(System %in% c("Lake", "River")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  ggplot() + 
  #unaligned
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #aligned
  stat_ecdf(aes(x = Conc, color = "Aligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = Conc, color = "Aligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #color
  scale_color_manual(name = "alignment",
                     values = vals) +
  #thresholds
  geom_vline(xintercept = Threshold1, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "1", color = "#00798c" , x = log10(Threshold1) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "2", color = "#66a182" , x = log10(Threshold2) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "3", color = '#edae49', x = log10(Threshold3) + 0.2, y = 0.1, size = 5)+
  geom_text(label = "4", color = "#d1495b", x = log10(Threshold4) + 0.2, y = 0.1, size = 5)+
  #aesthetics
   ylab("Cumulative Density") +
  xlab(paste0("Particles/L"))+
  scale_y_continuous(labels = scales::percent) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif) +
  theme.type +
  theme(legend.title = element_blank())

ECDF_freshwater
```

#### Manuscript Figure
```{r}
ECDF_freshwater_marine <- ggarrange(ECDF_marine, ECDF_freshwater,
                        labels = c("A", "B"), common.legend = TRUE,
                        legend = "bottom",
                        ncol = 2, nrow = 1)

plot(ECDF_freshwater_marine)

ggsave(plot =ECDF_freshwater_marine,
       filename = "ECDF_freshwater_marine_SI-Fig6.jpeg",
       path = "./Concentration data/Threshold_Manuscript_Figs/", 
       width = 11, height = 5, units = "in",
       dpi = 300)
```
# FRESHWATER + MARINE

Alvina wants the figure with both freshwater and marine occurence data in the same plot, with translocation and food dilution thresholds annotated in different plots.

## ECDF
### Food Dilution
```{r}
Threshold1_food = 0.5 
Threshold2_food = 8.1 
Threshold3_food = 14.1
Threshold4_food = 93.8
```

```{r}
vals <- c("Unaligned" = "black", "Aligned" = "blue")

ECDF_all_food <- globalData %>% 
  filter(sample.matrix.specific == "surface water") %>% 
 #  filter(System %in% c("Lake", "River")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  ggplot() + 
  #unaligned
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #aligned
  stat_ecdf(aes(x = Conc, color = "Aligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = Conc, color = "Aligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #color
  scale_color_manual(name = "alignment",
                     values = vals) +
  #thresholds
  geom_vline(xintercept = Threshold1_food, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2_food, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3_food, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4_food, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "1", color = "#00798c" , x = log10(Threshold1_food) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "2", color = "#66a182" , x = log10(Threshold2_food) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "3", color = '#edae49', x = log10(Threshold3_food) + 0.2, y = 0.1, size = 5)+
  geom_text(label = "4", color = "#d1495b", x = log10(Threshold4_food) + 0.2, y = 0.1, size = 5)+
  #aesthetics
   ylab("Cumulative Density") +
  xlab(paste0("Particles/L"))+
  scale_y_continuous(labels = scales::percent) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif) +
  theme.type +
  theme(legend.title = element_blank())

ECDF_all_food
```

### Translocation
```{r}
Threshold1_translocation = 236.34
Threshold2_translocation = 1312.43
Threshold3_translocation = 3096.59
Threshold4_translocation = 14397.3
```

```{r}
vals <- c("Unaligned" = "black", "Aligned" = "blue")

ECDF_all_translocation <- globalData %>% 
  filter(sample.matrix.specific == "surface water") %>% 
 #  filter(System %in% c("Lake", "River")) %>% 
  mutate(unaligned = particle.L.master / CF) %>%
  ggplot() + 
  #unaligned
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = unaligned, color = "Unaligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #aligned
  stat_ecdf(aes(x = Conc, color = "Aligned"),
            geom = "point", size = 1, alpha = 0.5) +
  stat_ecdf(aes(x = Conc, color = "Aligned"),geom = "step", linetype = 'solid', alpha = 0.3, size = 1) +
  #color
  scale_color_manual(name = "alignment",
                     values = vals) +
  #thresholds
  geom_vline(xintercept = Threshold1_translocation, linetype = 'dashed', color = "#00798c", size = 1.3) +
  geom_vline(xintercept = Threshold2_translocation, linetype = 'dashed', color = "#66a182" , size = 1.3) +
  geom_vline(xintercept = Threshold3_translocation, linetype = 'dashed', color = 	'#edae49', size = 1.3) +
  geom_vline(xintercept = Threshold4_translocation, linetype = 'dashed', color = 	"#d1495b", size = 1.3) +
  geom_text(label = "1", color = "#00798c" , x = log10(Threshold1_translocation) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "2", color = "#66a182" , x = log10(Threshold2_translocation) - 0.2, y = 0.1, size = 5)+
  geom_text(label = "3", color = '#edae49', x = log10(Threshold3_translocation) + 0.2, y = 0.1, size = 5)+
  geom_text(label = "4", color = "#d1495b", x = log10(Threshold4_translocation) + 0.2, y = 0.1, size = 5)+
  #aesthetics
   ylab("Cumulative Density") +
  xlab(paste0("Particles/L"))+
  scale_y_continuous(labels = scales::percent) +
  scale_x_log10(breaks = scales::trans_breaks("log10", function(x) 10^x),labels = comma_signif) +
  theme.type +
  theme(legend.title = element_blank())

ECDF_all_translocation
```

#### Manuscript Figure
```{r}
ECDF_all_ERM <- ggarrange(ECDF_all_food, ECDF_all_translocation,
                        labels = c("A", "B"), common.legend = TRUE,
                        legend = "bottom",
                        ncol = 2, nrow = 1)

plot(ECDF_all_ERM)

ggsave(plot =ECDF_all_ERM,
       filename = "ECDF_all_ERM_SI-Fig6.jpeg",
       path = "./Concentration data/Threshold_Manuscript_Figs/", 
       width = 11, height = 5, units = "in",
       dpi = 300)
```