---
title: "SSD Sensitivity Analysis"
author: "Scott Coffin"
date: "3/10/2021"
output:   
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
    includes:
     # after_body: footer.html
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
```

```{r libraries, include=FALSE}
library(tidyverse)
library(calecopal)
library(ssdtools)
library(DT)
library(plotly)
library(gridExtra)
library(grid)
library(wesanderson)
library(ggdark)
library(broom)
library(knitr)
library(viridis)
library(ggrepel)
library(scales)
library(gt)
```

```{r parameters, include=FALSE}
# Paramaters
x2D_set = 5000#100 #upper size range (microns)
x1D_set = 100#1 #lower size range (microns)
particle_mass_check_ssd <- 'mg/L' #assign whether or not to use particles/mL or mass/mL
```

```{r Theme, include=FALSE}
#Theme type
     theme.type<- theme_gray(base_size = 20)#,
                  #     dark_theme_bw(base_size = 20)
     #color selection
     fill.type <-    scale_fill_viridis(discrete = TRUE)#,
                         #scale_fill_brewer(palette = "Paired"),
                         # scale_fill_tron(),
                         # scale_fill_locuszoom(),
                         # scale_fill_d3(),
                         # scale_fill_npg(),
                         # scale_fill_jama())
     #color selection
     color.type <- scale_color_viridis(discrete = TRUE)#,
                         # scale_color_brewer(palette = "Paired"),
                         # scale_color_tron(),
                         # scale_color_locuszoom(),
                         # scale_color_d3(),
                         # scale_color_npg(),
                         # scale_color_jama())

```

```{r import, include=FALSE}
# Data Import
#require(readr)
#load aoc_z into dataframe. This file is generated from RDA_Maker.R
#aoc_z <- readRDS(file = "Tox Data/aoc_z.Rda")
source("Tox Data/RDA_Maker.R")
```


# Data Collapse

```{r data collapse, include=FALSE}
#### filter by tier first ###
#Tier 1 is HC5 for all endpoints
tier1 <- aoc_z %>% 
  filter(environment != c("Terrestrial", "Not Reported"),
         acute.chronic_f != "Unavailable",
         tier_zero_tech_f == "Red Criteria Passed",
         tier_zero_risk_f == "Red Criteria Passed",
         shape_f != "Not Reported",
         lvl1_f == "Fitness") %>% 
  replace_na(list(af.time = 1)) %>% 
  mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  #convert acute to chronic and NOEC to LOEC
  drop_na(dose_new)

#Tier 2 is HC5 for all tissue and above endpoints
tier2 <- aoc_z %>% 
  filter(environment != c("Terrestrial", "Not Reported"),
         acute.chronic_f != "Unavailable",
         tier_zero_tech_f == "Red Criteria Passed",
         tier_zero_risk_f == "Red Criteria Passed",
         shape_f != "Not Reported",
         lvl1_f == "Fitness",
         bio_f %in% c("Tissue","Organism","Population")) %>% 
  replace_na(list(af.time = 1)) %>% 
  mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  #convert acute to chronic and NOEC to LOEC
  drop_na(dose_new)

#Tiers 3 and 4 are for organism level endpoints only
tier3.4 <- aoc_z %>% 
  filter(environment != c("Terrestrial", "Not Reported"),
         acute.chronic_f != "Unavailable",
         tier_zero_tech_f == "Red Criteria Passed",
         tier_zero_risk_f == "Red Criteria Passed",
         shape_f != "Not Reported",
         lvl1_f == "Fitness",
         bio_f == "Organism") %>% 
  replace_na(list(af.time = 1)) %>% 
  mutate(dose_new = dose.mg.L.master / (af.time * af.noec)) %>%  #convert acute to chronic and NOEC to LOEC
  drop_na(dose_new)

#### Size filters ####
### 100 - 5000 um
tier1_large <- tier1 %>% 
  filter(size_f %in% c("10µm < 100µm", "100µm < 1mm", "1mm < 5mm"))

tier2_large <- tier2 %>% 
  filter(size_f %in% c("10µm < 100µm", "100µm < 1mm", "1mm < 5mm"))

tier3.4_large <- tier3.4 %>% 
  filter(size_f %in% c("10µm < 100µm", "100µm < 1mm", "1mm < 5mm"))

### 10 - 100 um
tier1_small <- tier1 %>% 
  filter(size_f %in% c("1µm < 10µm"))

tier2_small <- tier2 %>% 
  filter(size_f %in% c("1µm < 10µm"))

tier3.4_small <- tier3.4 %>% 
  filter(size_f %in% c("1µm < 10µm"))

#### Data Collapsing ####

## Tier 1 ##
# tier 1 large
tier1_large_median <- tier1_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier1_large_mean <- tier1_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier1_large_1q <- tier1_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25))

tier1_large_min <- tier1_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))

# tier 1 small
tier1_small_median <- tier1_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier1_small_mean <- tier1_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier1_small_1q <- tier1_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25))

tier1_small_min <- tier1_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))

## Tier 2 ##
# tier 1 large
tier2_large_median <- tier2_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier2_large_mean <- tier2_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier2_large_1q <- tier2_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25))

tier2_large_min <- tier1_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))

# tier 1 small
tier2_small_median <- tier2_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier2_small_mean <- tier2_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier2_small_1q <- tier2_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25))

tier2_small_min <- tier2_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))

## Tier 3 and 4 ##
# tier 3 large
tier3.4_large_median <- tier3.4_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier3.4_large_mean <- tier3.4_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier3.4_large_1q <- tier3.4_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25))

tier3.4_large_min <- tier3.4_large %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))

# tier 3 small
tier3.4_small_median <- tier3.4_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = median(dose_new))

tier3.4_small_mean <- tier3.4_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = mean(dose_new))

tier3.4_small_1q <- tier3.4_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = quantile(dose_new, 0.25), na.rm = TRUE)

tier3.4_small_min <- tier3.4_small %>% 
  group_by(Species, Group) %>% 
  summarize(Conc = min(dose_new))
```

# SSD Build
## Tier 1
### Large 
#### Tier1_large_median

```{r include=FALSE}
###### --modelling ####
dists_tier1_large_median <- ssd_fit_dists(tier1_large_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_large_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_large_median <- as.data.frame(ssd_gof(dists_tier1_large_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_large_median <- predict(dists_tier1_large_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_large_median <- tier1_large_median[order(tier1_large_median$Conc), ]
SSD_tier1_large_median$frac <- ppoints(tier1_large_median$Conc, 0.5)

tier1_large_median_hc5 <- c(pred_tier1_large_median$est[5])
pred_tier1_large_median$est_format <-format(pred_tier1_large_median$est, digits = 3, scientific = TRUE)
```


```{r}
tier1_large_median_ggplot <- ggplot(pred_tier1_large_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_large_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_large_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_large_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier1_large_median_ggplot
```

#### tier1_large_mean

```{r include=FALSE}
###### --modelling ####
dists_tier1_large_mean <- ssd_fit_dists(tier1_large_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_large_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_large_mean <- as.data.frame(ssd_gof(dists_tier1_large_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_large_mean <- predict(dists_tier1_large_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_large_mean <- tier1_large_mean[order(tier1_large_mean$Conc), ]
SSD_tier1_large_mean$frac <- ppoints(tier1_large_mean$Conc, 0.5)

tier1_large_mean_hc5 <- c(pred_tier1_large_mean$est[5])
pred_tier1_large_mean$est_format <-format(pred_tier1_large_mean$est, digits = 3, scientific = TRUE)
```

```{r}
tier1_large_mean_ggplot <- ggplot(pred_tier1_large_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_large_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_large_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_large_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier1_large_mean_ggplot
```

#### tier1_large_1q
```{r include=FALSE}
###### --modelling ####
dists_tier1_large_1q <- ssd_fit_dists(tier1_large_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_large_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_large_1q <- as.data.frame(ssd_gof(dists_tier1_large_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_large_1q <- predict(dists_tier1_large_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_large_1q <- tier1_large_1q[order(tier1_large_1q$Conc), ]
SSD_tier1_large_1q$frac <- ppoints(tier1_large_1q$Conc, 0.5)

tier1_large_1q_hc5 <- c(pred_tier1_large_1q$est[5])
pred_tier1_large_1q$est_format <-format(pred_tier1_large_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier1_large_1q_ggplot <- ggplot(pred_tier1_large_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_large_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_large_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_large_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier1_large_1q_ggplot
```

#### tier1_large_min
```{r include=FALSE}
###### --modelling ####
dists_tier1_large_min <- ssd_fit_dists(tier1_large_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_large_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_large_min <- as.data.frame(ssd_gof(dists_tier1_large_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_large_min <- predict(dists_tier1_large_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_large_min <- tier1_large_min[order(tier1_large_min$Conc), ]
SSD_tier1_large_min$frac <- ppoints(tier1_large_min$Conc, 0.5)

tier1_large_min_hc5 <- c(pred_tier1_large_min$est[5])
pred_tier1_large_min$est_format <-format(pred_tier1_large_min$est, digits = 3, scientific = TRUE)

#ggplot
tier1_large_min_ggplot <- ggplot(pred_tier1_large_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_large_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_large_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_large_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier1_large_min_ggplot
```

### small 
#### Tier1_small_median

```{r include=FALSE}
###### --modelling ####
dists_tier1_small_median <- ssd_fit_dists(tier1_small_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_small_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_small_median <- as.data.frame(ssd_gof(dists_tier1_small_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_small_median <- predict(dists_tier1_small_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_small_median <- tier1_small_median[order(tier1_small_median$Conc), ]
SSD_tier1_small_median$frac <- ppoints(tier1_small_median$Conc, 0.5)

tier1_small_median_hc5 <- c(pred_tier1_small_median$est[5])
pred_tier1_small_median$est_format <-format(pred_tier1_small_median$est, digits = 3, scientific = TRUE)

#ggplot
tier1_small_median_ggplot <- ggplot(pred_tier1_small_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_small_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_small_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_small_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier1_small_median_ggplot
```

#### tier1_small_mean

```{r include=FALSE}
###### --modelling ####
dists_tier1_small_mean <- ssd_fit_dists(tier1_small_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_small_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_small_mean <- as.data.frame(ssd_gof(dists_tier1_small_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_small_mean <- predict(dists_tier1_small_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_small_mean <- tier1_small_mean[order(tier1_small_mean$Conc), ]
SSD_tier1_small_mean$frac <- ppoints(tier1_small_mean$Conc, 0.5)

tier1_small_mean_hc5 <- c(pred_tier1_small_mean$est[5])
pred_tier1_small_mean$est_format <-format(pred_tier1_small_mean$est, digits = 3, scientific = TRUE)

## ggplot
tier1_small_mean_ggplot <- ggplot(pred_tier1_small_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_small_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_small_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_small_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier1_small_mean_ggplot
```

#### tier1_small_1q
```{r include=FALSE}
###### --modelling ####
dists_tier1_small_1q <- ssd_fit_dists(tier1_small_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_small_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_small_1q <- as.data.frame(ssd_gof(dists_tier1_small_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_small_1q <- predict(dists_tier1_small_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_small_1q <- tier1_small_1q[order(tier1_small_1q$Conc), ]
SSD_tier1_small_1q$frac <- ppoints(tier1_small_1q$Conc, 0.5)

tier1_small_1q_hc5 <- c(pred_tier1_small_1q$est[5])
pred_tier1_small_1q$est_format <-format(pred_tier1_small_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier1_small_1q_ggplot <- ggplot(pred_tier1_small_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_small_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_small_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_small_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier1_small_1q_ggplot
```

#### tier1_small_min
```{r include=FALSE}
###### --modelling ####
dists_tier1_small_min <- ssd_fit_dists(tier1_small_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier1_small_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier1_small_min <- as.data.frame(ssd_gof(dists_tier1_small_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier1_small_min <- predict(dists_tier1_small_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier1_small_min <- tier1_small_min[order(tier1_small_min$Conc), ]
SSD_tier1_small_min$frac <- ppoints(tier1_small_min$Conc, 0.5)

tier1_small_min_hc5 <- c(pred_tier1_small_min$est[5])
pred_tier1_small_min$est_format <-format(pred_tier1_small_min$est, digits = 3, scientific = TRUE)

#ggplot
tier1_small_min_ggplot <- ggplot(pred_tier1_small_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier1_small_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier1_small_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier1_small_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier1_small_min_ggplot
```

## tier 2
### Large 
#### tier2_large_median

```{r include=FALSE}
###### --modelling ####
dists_tier2_large_median <- ssd_fit_dists(tier2_large_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_large_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_large_median <- as.data.frame(ssd_gof(dists_tier2_large_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_large_median <- predict(dists_tier2_large_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_large_median <- tier2_large_median[order(tier2_large_median$Conc), ]
SSD_tier2_large_median$frac <- ppoints(tier2_large_median$Conc, 0.5)

tier2_large_median_hc5 <- c(pred_tier2_large_median$est[5])
pred_tier2_large_median$est_format <-format(pred_tier2_large_median$est, digits = 3, scientific = TRUE)
```


```{r}
tier2_large_median_ggplot <- ggplot(pred_tier2_large_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_large_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_large_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_large_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_large_median_ggplot
```

#### tier2_large_mean

```{r include=FALSE}
###### --modelling ####
dists_tier2_large_mean <- ssd_fit_dists(tier2_large_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_large_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_large_mean <- as.data.frame(ssd_gof(dists_tier2_large_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_large_mean <- predict(dists_tier2_large_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_large_mean <- tier2_large_mean[order(tier2_large_mean$Conc), ]
SSD_tier2_large_mean$frac <- ppoints(tier2_large_mean$Conc, 0.5)

tier2_large_mean_hc5 <- c(pred_tier2_large_mean$est[5])
pred_tier2_large_mean$est_format <-format(pred_tier2_large_mean$est, digits = 3, scientific = TRUE)
```

```{r}
tier2_large_mean_ggplot <- ggplot(pred_tier2_large_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_large_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_large_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_large_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_large_mean_ggplot
```

#### tier2_large_1q
```{r include=FALSE}
###### --modelling ####
dists_tier2_large_1q <- ssd_fit_dists(tier2_large_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_large_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_large_1q <- as.data.frame(ssd_gof(dists_tier2_large_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_large_1q <- predict(dists_tier2_large_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_large_1q <- tier2_large_1q[order(tier2_large_1q$Conc), ]
SSD_tier2_large_1q$frac <- ppoints(tier2_large_1q$Conc, 0.5)

tier2_large_1q_hc5 <- c(pred_tier2_large_1q$est[5])
pred_tier2_large_1q$est_format <-format(pred_tier2_large_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier2_large_1q_ggplot <- ggplot(pred_tier2_large_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_large_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_large_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_large_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_large_1q_ggplot
```

#### tier2_large_min
```{r include=FALSE}
###### --modelling ####
dists_tier2_large_min <- ssd_fit_dists(tier2_large_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_large_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_large_min <- as.data.frame(ssd_gof(dists_tier2_large_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_large_min <- predict(dists_tier2_large_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_large_min <- tier2_large_min[order(tier2_large_min$Conc), ]
SSD_tier2_large_min$frac <- ppoints(tier2_large_min$Conc, 0.5)

tier2_large_min_hc5 <- c(pred_tier2_large_min$est[5])
pred_tier2_large_min$est_format <-format(pred_tier2_large_min$est, digits = 3, scientific = TRUE)

#ggplot
tier2_large_min_ggplot <- ggplot(pred_tier2_large_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_large_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_large_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_large_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_large_min_ggplot
```
### small 
#### tier2_small_median

```{r include=FALSE}
###### --modelling ####
dists_tier2_small_median <- ssd_fit_dists(tier2_small_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_small_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_small_median <- as.data.frame(ssd_gof(dists_tier2_small_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_small_median <- predict(dists_tier2_small_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_small_median <- tier2_small_median[order(tier2_small_median$Conc), ]
SSD_tier2_small_median$frac <- ppoints(tier2_small_median$Conc, 0.5)

tier2_small_median_hc5 <- c(pred_tier2_small_median$est[5])
pred_tier2_small_median$est_format <-format(pred_tier2_small_median$est, digits = 3, scientific = TRUE)

#ggplot
tier2_small_median_ggplot <- ggplot(pred_tier2_small_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_small_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_small_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_small_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_small_median_ggplot
```

#### tier2_small_mean

```{r include=FALSE}
###### --modelling ####
dists_tier2_small_mean <- ssd_fit_dists(tier2_small_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_small_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_small_mean <- as.data.frame(ssd_gof(dists_tier2_small_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_small_mean <- predict(dists_tier2_small_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_small_mean <- tier2_small_mean[order(tier2_small_mean$Conc), ]
SSD_tier2_small_mean$frac <- ppoints(tier2_small_mean$Conc, 0.5)

tier2_small_mean_hc5 <- c(pred_tier2_small_mean$est[5])
pred_tier2_small_mean$est_format <-format(pred_tier2_small_mean$est, digits = 3, scientific = TRUE)

## ggplot
tier2_small_mean_ggplot <- ggplot(pred_tier2_small_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_small_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_small_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_small_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier2_small_mean_ggplot
```

#### tier2_small_1q
```{r include=FALSE}
###### --modelling ####
dists_tier2_small_1q <- ssd_fit_dists(tier2_small_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_small_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_small_1q <- as.data.frame(ssd_gof(dists_tier2_small_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_small_1q <- predict(dists_tier2_small_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_small_1q <- tier2_small_1q[order(tier2_small_1q$Conc), ]
SSD_tier2_small_1q$frac <- ppoints(tier2_small_1q$Conc, 0.5)

tier2_small_1q_hc5 <- c(pred_tier2_small_1q$est[5])
pred_tier2_small_1q$est_format <-format(pred_tier2_small_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier2_small_1q_ggplot <- ggplot(pred_tier2_small_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_small_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_small_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_small_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier2_small_1q_ggplot
```

#### tier2_small_min
```{r include=FALSE}
###### --modelling ####
dists_tier2_small_min <- ssd_fit_dists(tier2_small_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier2_small_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier2_small_min <- as.data.frame(ssd_gof(dists_tier2_small_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier2_small_min <- predict(dists_tier2_small_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier2_small_min <- tier2_small_min[order(tier2_small_min$Conc), ]
SSD_tier2_small_min$frac <- ppoints(tier2_small_min$Conc, 0.5)

tier2_small_min_hc5 <- c(pred_tier2_small_min$est[5])
pred_tier2_small_min$est_format <-format(pred_tier2_small_min$est, digits = 3, scientific = TRUE)

#ggplot
tier2_small_min_ggplot <- ggplot(pred_tier2_small_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier2_small_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier2_small_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier2_small_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier2_small_min_ggplot
```

## Tier 3
### Large 
#### tier3.4_large_median

```{r include=FALSE}
###### --modelling ####
dists_tier3.4_large_median <- ssd_fit_dists(tier3.4_large_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_large_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_large_median <- as.data.frame(ssd_gof(dists_tier3.4_large_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_large_median <- predict(dists_tier3.4_large_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_large_median <- tier3.4_large_median[order(tier3.4_large_median$Conc), ]
SSD_tier3.4_large_median$frac <- ppoints(tier3.4_large_median$Conc, 0.5)

tier3.4_large_median_hc5 <- c(pred_tier3.4_large_median$est[5])
#extrac hc10 (tier 4)
tier3.4_large_median_hc10 <- c(pred_tier3.4_large_median$est[10])
pred_tier3.4_large_median$est_format <-format(pred_tier3.4_large_median$est, digits = 3, scientific = TRUE)
```


```{r}
tier3.4_large_median_ggplot <- ggplot(pred_tier3.4_large_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_large_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_large_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_large_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_large_median_ggplot
```

#### tier3.4_large_mean

```{r include=FALSE}
###### --modelling ####
dists_tier3.4_large_mean <- ssd_fit_dists(tier3.4_large_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_large_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_large_mean <- as.data.frame(ssd_gof(dists_tier3.4_large_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_large_mean <- predict(dists_tier3.4_large_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_large_mean <- tier3.4_large_mean[order(tier3.4_large_mean$Conc), ]
SSD_tier3.4_large_mean$frac <- ppoints(tier3.4_large_mean$Conc, 0.5)

tier3.4_large_mean_hc5 <- c(pred_tier3.4_large_mean$est[5])
#extrac hc10 (tier 4)
tier3.4_large_mean_hc10 <- c(pred_tier3.4_large_mean$est[10])

pred_tier3.4_large_mean$est_format <-format(pred_tier3.4_large_mean$est, digits = 3, scientific = TRUE)
```

```{r}
tier3.4_large_mean_ggplot <- ggplot(pred_tier3.4_large_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_large_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_large_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_large_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_large_mean_ggplot
```

#### tier3.4_large_1q
```{r include=FALSE}
###### --modelling ####
dists_tier3.4_large_1q <- ssd_fit_dists(tier3.4_large_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_large_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_large_1q <- as.data.frame(ssd_gof(dists_tier3.4_large_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_large_1q <- predict(dists_tier3.4_large_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_large_1q <- tier3.4_large_1q[order(tier3.4_large_1q$Conc), ]
SSD_tier3.4_large_1q$frac <- ppoints(tier3.4_large_1q$Conc, 0.5)

tier3.4_large_1q_hc5 <- c(pred_tier3.4_large_1q$est[5])
#extrac hc10 (tier 4)
tier3.4_large_1q_hc10 <- c(pred_tier3.4_large_1q$est[10])
pred_tier3.4_large_1q$est_format <-format(pred_tier3.4_large_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier3.4_large_1q_ggplot <- ggplot(pred_tier3.4_large_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_large_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_large_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_large_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_large_1q_ggplot
```

#### tier3.4_large_min
```{r include=FALSE}
###### --modelling ####
dists_tier3.4_large_min <- ssd_fit_dists(tier3.4_large_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_large_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_large_min <- as.data.frame(ssd_gof(dists_tier3.4_large_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_large_min <- predict(dists_tier3.4_large_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_large_min <- tier3.4_large_min[order(tier3.4_large_min$Conc), ]
SSD_tier3.4_large_min$frac <- ppoints(tier3.4_large_min$Conc, 0.5)

tier3.4_large_min_hc5 <- c(pred_tier3.4_large_min$est[5])
#extrac hc10 (tier 4)
tier3.4_large_min_hc10 <- c(pred_tier3.4_large_min$est[10])
pred_tier3.4_large_min$est_format <-format(pred_tier3.4_large_min$est, digits = 3, scientific = TRUE)

#ggplot
tier3.4_large_min_ggplot <- ggplot(pred_tier3.4_large_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_large_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_large_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_large_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_large_min_ggplot
```

### small 
#### tier3.4_small_median

```{r include=FALSE}
###### --modelling ####
dists_tier3.4_small_median <- ssd_fit_dists(tier3.4_small_median, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_small_median) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_small_median <- as.data.frame(ssd_gof(dists_tier3.4_small_median)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_small_median <- predict(dists_tier3.4_small_median, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_small_median <- tier3.4_small_median[order(tier3.4_small_median$Conc), ]
SSD_tier3.4_small_median$frac <- ppoints(tier3.4_small_median$Conc, 0.5)

tier3.4_small_median_hc5 <- c(pred_tier3.4_small_median$est[5])
#extrac hc10 (tier 4)
tier3.4_small_median_hc10 <- c(pred_tier3.4_small_median$est[10])

pred_tier3.4_small_median$est_format <-format(pred_tier3.4_small_median$est, digits = 3, scientific = TRUE)

#ggplot
tier3.4_small_median_ggplot <- ggplot(pred_tier3.4_small_median,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_small_median,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_small_median, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_small_median") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_small_median_ggplot
```

#### tier3.4_small_mean

```{r include=FALSE}
###### --modelling ####
dists_tier3.4_small_mean <- ssd_fit_dists(tier3.4_small_mean, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_small_mean) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_small_mean <- as.data.frame(ssd_gof(dists_tier3.4_small_mean)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_small_mean <- predict(dists_tier3.4_small_mean, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_small_mean <- tier3.4_small_mean[order(tier3.4_small_mean$Conc), ]
SSD_tier3.4_small_mean$frac <- ppoints(tier3.4_small_mean$Conc, 0.5)

tier3.4_small_mean_hc5 <- c(pred_tier3.4_small_mean$est[5])
#extrac hc10 (tier 4)
tier3.4_small_mean_hc10 <- c(pred_tier3.4_small_mean$est[10])

pred_tier3.4_small_mean$est_format <-format(pred_tier3.4_small_mean$est, digits = 3, scientific = TRUE)

## ggplot
tier3.4_small_mean_ggplot <- ggplot(pred_tier3.4_small_mean,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_small_mean,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_small_mean, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_small_mean") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme
tier3.4_small_mean_ggplot
```

#### tier3.4_small_1q
```{r include=FALSE}
###### --modelling ####
dists_tier3.4_small_1q <- ssd_fit_dists(tier3.4_small_1q, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_small_1q) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_small_1q <- as.data.frame(ssd_gof(dists_tier3.4_small_1q)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_small_1q <- predict(dists_tier3.4_small_1q, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_small_1q <- tier3.4_small_1q[order(tier3.4_small_1q$Conc), ]
SSD_tier3.4_small_1q$frac <- ppoints(tier3.4_small_1q$Conc, 0.5)

tier3.4_small_1q_hc5 <- c(pred_tier3.4_small_1q$est[5])
#extrac hc10 (tier 4)
tier3.4_small_1q_hc10 <- c(pred_tier3.4_small_1q$est[10])

pred_tier3.4_small_1q$est_format <-format(pred_tier3.4_small_1q$est, digits = 3, scientific = TRUE)

#ggplot
tier3.4_small_1q_ggplot <- ggplot(pred_tier3.4_small_1q,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_small_1q,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_small_1q, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_small_1q") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier3.4_small_1q_ggplot
```

#### tier3.4_small_min
```{r include=FALSE}
###### --modelling ####
dists_tier3.4_small_min <- ssd_fit_dists(tier3.4_small_min, left = "Conc", dists = c("weibull", "llogis", "lnorm", "gamma", "lgumbel"), computable = FALSE, silent = FALSE) 

#autoplot(dists_tier3.4_small_min) #plots the distribution in ggplotier2
#### Goodness of fit ####
gof_tier3.4_small_min <- as.data.frame(ssd_gof(dists_tier3.4_small_min)) %>% mutate_if(is.numeric, ~ signif(., 3))

#### Prediction ####
set.seed(99)
pred_tier3.4_small_min <- predict(dists_tier3.4_small_min, average = TRUE, ic = "aicc", nboot = 10, ci= TRUE) 

#order data
SSD_tier3.4_small_min <- tier3.4_small_min[order(tier3.4_small_min$Conc), ]
SSD_tier3.4_small_min$frac <- ppoints(tier3.4_small_min$Conc, 0.5)

#extract hc5
tier3.4_small_min_hc5 <- c(pred_tier3.4_small_min$est[5])

#extrac hc10 (tier 4)
tier3.4_small_min_hc10 <- c(pred_tier3.4_small_min$est[10])
pred_tier3.4_small_min$est_format <-format(pred_tier3.4_small_min$est, digits = 3, scientific = TRUE)

#ggplot
tier3.4_small_min_ggplot <- ggplot(pred_tier3.4_small_min,aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, color = "blue", fill = "lightblue") +
      geom_line(aes_string(y = "percent/100"), color = "gray") +
      geom_point(data = SSD_tier3.4_small_min,aes(x = Conc, y =frac, color = Group)) + 
      geom_text_repel(data = SSD_tier3.4_small_min, aes(x = Conc, y = frac, label = Species, color = Group), nudge_x = 0.2, size = 3, segment.alpha = 0.5) + 
      scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) + 
  labs(title = "tier3.4_small_min") +
      fill.type + #user-selected
      color.type + #user-selected
      theme.type #user theme

tier3.4_small_min_ggplot
```

# Combine plots
## Tier 1
### Smalls
```{r}
### tier1_small_median
ggplot(pred_tier1_small_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier1_small_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier1_small_mean ###
  geom_xribbon(data = pred_tier1_small_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier1_small_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier1_small_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier1_small_1q ###
    geom_xribbon(data = pred_tier1_small_1q, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier1_small_1q, aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier1_small_1q, aes(x = Conc, y =frac), 
                 color = "green") +  
  
### fourth line: tier1_small_min ###
    geom_xribbon(data = pred_tier1_small_min, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier1_small_min, aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier1_small_min, aes(x = Conc, y =frac), 
                 color = "grey10") +  
### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 1)",
       subtitle = ("Size Range: 1 - 10 um; Red = median, blue = mean, green = first quartile, grey = minimum"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```
### larges
```{r}
### tier1_large_median
ggplot(pred_tier1_large_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier1_large_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier1_large_mean ###
  geom_xribbon(data = pred_tier1_large_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier1_large_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier1_large_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier1_large_1q ###
    geom_xribbon(data = pred_tier1_large_1q, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier1_large_1q, aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier1_large_1q, aes(x = Conc, y =frac), 
                 color = "green") +  
  
  ### fourth line: tier1_large_min ###
    geom_xribbon(data = pred_tier1_large_min, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier1_large_min, aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier1_large_min, aes(x = Conc, y =frac), 
                 color = "grey10") +  
### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 1)",
       subtitle = ("Size Range: 10 - 5,000 um; Red = median, blue = mean, green = first quartile, grey = minimum"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

## Tier 2
### Smalls
```{r}
### tier2_small_median
ggplot(pred_tier2_small_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier2_small_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier2_small_mean ###
  geom_xribbon(data = pred_tier2_small_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier2_small_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier2_small_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier2_small_1q ###
    geom_xribbon(data = pred_tier2_small_1q, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier2_small_1q, aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier2_small_1q, aes(x = Conc, y =frac), 
                 color = "green") +  
  ### fourth line: tier2_small_min  ###
    geom_xribbon(data = pred_tier2_small_min , aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier2_small_min , aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier2_small_min , aes(x = Conc, y =frac), 
                 color = "grey10") +  
  ### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 2)",
       subtitle = ("Size Range: 1 - 10 um; Red = median, blue = mean, green = first quartile, grey = minimum"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```
### larges
```{r}
### tier2_large_median
ggplot(pred_tier2_large_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier2_large_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier2_large_mean ###
  geom_xribbon(data = pred_tier2_large_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier2_large_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier2_large_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier2_large_1q  ###
    geom_xribbon(data = pred_tier2_large_1q , aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier2_large_1q , aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier2_large_1q , aes(x = Conc, y =frac), 
                 color = "green") +  
  
    ### fourth line: tier2_large_min  ###
    geom_xribbon(data = pred_tier2_large_min , aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier2_large_min , aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier2_large_min , aes(x = Conc, y =frac), 
                 color = "grey10") +  
### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 2)",
       subtitle = ("Size Range: 10 - 5,000 um; Red = median, blue = mean, green = first quartile, grey = minimum"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

## Tier 3/4
### Smalls
```{r}
### tier3.4_small_median
ggplot(pred_tier3.4_small_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier3.4_small_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier3.4_small_mean ###
  geom_xribbon(data = pred_tier3.4_small_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier3.4_small_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier3.4_small_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier3.4_small_1q ###
    geom_xribbon(data = pred_tier3.4_small_1q, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier3.4_small_1q, aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier3.4_small_1q, aes(x = Conc, y =frac), 
                 color = "green") +  
  
      ### fourth line: tier3.4_small_min ###
    geom_xribbon(data = pred_tier3.4_small_min, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier3.4_small_min, aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier3.4_small_min, aes(x = Conc, y =frac), 
                 color = "grey10") + 
### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 3/4)",
       subtitle = ("Size Range: 1 - 10 um; Red = median, blue = mean, green = first quartile, grey = min"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```
### larges
```{r}
### tier3.4_large_median
ggplot(pred_tier3.4_large_median, aes_string(x = "est")) +
      geom_xribbon(aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
                   color = "red", fill = "red") +
      geom_line(aes_string(y = "percent/100"), 
                color = "red") +
      geom_point(data = SSD_tier3.4_large_median,aes(x = Conc, y =frac), 
                 color = "red") + 
      
### second line: tier3.4_large_mean ###
  geom_xribbon(data = pred_tier3.4_large_mean, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "blue", fill = "blue") +
      geom_line(data = pred_tier3.4_large_mean, aes_string(y = "percent/100"), 
                color = "blue") +
      geom_point(data = SSD_tier3.4_large_mean, aes(x = Conc, y =frac), 
                 color = "blue") +  
### third line: tier3.4_large_1q ###
    geom_xribbon(data = pred_tier3.4_large_1q, aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "green", fill = "green") +
      geom_line(data = pred_tier3.4_large_1q, aes_string(y = "percent/100"), 
                color = "green") +
      geom_point(data = SSD_tier3.4_large_1q, aes(x = Conc, y =frac), 
                 color = "green") +  
  
      ### fourth line: tier3.4_large_min  ###
    geom_xribbon(data = pred_tier3.4_large_min , aes_string(xmin = "lcl", xmax = "ucl", y = "percent/100"), alpha = 0.2, 
               color = "grey10", fill = "grey10") +
      geom_line(data = pred_tier3.4_large_min , aes_string(y = "percent/100"), 
                color = "grey10") +
      geom_point(data = SSD_tier3.4_large_min , aes(x = Conc, y =frac), 
                 color = "grey10") +  
### plot parameters ###
  scale_y_continuous("Species Affected (%)", labels = scales::percent, limits = c(0,1)) +
      xlab(particle_mass_check_ssd)+
      coord_trans(x = "log10") +
      scale_x_continuous(breaks = scales::trans_breaks("log10", function(x) 10^x, n = 15),
                         labels = trans_format("log10", scales::math_format(10^.x))) +
  labs(title = "Species Sensitivity Distributions by Data Collapsing (Tier 3/4)",
       subtitle = ("Size Range: 10 - 5,000 um; Red = median, blue = mean, green = first quartile, grey = minimum"),
       caption = ("All endpoints")) +
  theme.type +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 20),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

# Summary Table
```{r}
# Small/Large x tier 1/2/3 x median/mean/q1 = 2 x 3 x 3 = 18 comparisons
larges <- list(tier1_large_median_hc5, tier1_large_mean_hc5, tier1_large_1q_hc5, tier1_large_min_hc5,
          #tier 2
          tier2_large_median_hc5, tier2_large_mean_hc5, tier2_large_1q_hc5, tier2_large_min_hc5,
          #tier 3 (hc5)
          tier3.4_large_median_hc5, tier3.4_large_mean_hc5, tier3.4_large_1q_hc5, tier3.4_large_min_hc5,
          #tier 4 (hc10)
          tier3.4_large_median_hc10, tier3.4_large_mean_hc10, tier3.4_large_1q_hc10, tier3.4_large_min_hc10)

smalls <- list(tier1_small_median_hc5, tier1_small_mean_hc5, tier1_small_1q_hc5, tier1_small_min_hc5,
          #tier 2
          tier2_small_median_hc5, tier2_small_mean_hc5, tier2_small_1q_hc5, tier2_small_min_hc5,
          #tier 3.4
          tier3.4_small_median_hc5, tier3.4_small_mean_hc5, tier3.4_small_1q_hc5, tier3.4_small_min_hc5,
          #tier 4 (hc10)
          tier3.4_small_median_hc10, tier3.4_small_mean_hc10, tier3.4_small_1q_hc10, tier3.4_small_min_hc10)

## Conver to data frames
largesdf <- data.frame(t(matrix(unlist(larges), ncol = 4)))
smallsdf <- data.frame(t(matrix(unlist(smalls), ncol = 4))) #t = transpose

# convert mg/L to ug/L
largesdf <- largesdf * 1000
smallsdf <- smallsdf * 1000

# rename columns and rows
colnames(largesdf) <- c("Median", "Mean", "Quartile_1", "Minimum")
colnames(smallsdf) <- c("Median", "Mean", "Quartile_1", "Minimum")

rownames(largesdf) <- c("Tier 1", "Tier 2", "Tier 3", "Tier 4")
rownames(smallsdf) <- c("Tier 1", "Tier 2", "Tier 3", "Tier 4")

# provide tier column
largestibble <- tibble(largesdf,
                       Tier = rownames(largesdf))

#re-arrange columns
largestibble <- largestibble[c("Tier", "Mean", "Median", "Quartile_1", "Minimum")]

# provide tier column
smallstibble <- tibble(smallsdf,
                       Tier = rownames(smallsdf))

#re-arrange columns
smallstibble <- smallstibble[c("Tier", "Mean", "Median", "Quartile_1", "Minimum")]

#function to create tables
hp_table <- function(x){
  #get full value range for whole table conditional formatting
  full_val_range <- x %>% 
    ungroup %>% 
    select_if(is.numeric) %>% 
    range
  
  #format colors accordingly
  gt(x) %>% 
    fmt_number(columns = vars(Mean, Median, Quartile_1, Minimum),
             n_sigfig = 2,
             use_seps = TRUE) %>% 
    data_color(
    columns = names(x)[2:5],
    colors = scales::col_numeric(
      palette = paletteer::paletteer_d(palette = "ggsci::teal_material") %>% as.character(),domain = full_val_range),
    alpha = 0.75) %>% 
    tab_options(column_labels.hidden = FALSE) %>% 
    as_raw_html() # return as html
}

#larges gt
larges_gt <- largestibble %>% 
  hp_table()

#smalls gt
smalls_gt <- smallstibble %>% 
  hp_table() 
  
## Combine
data.tables <- data.frame(smalls_table = smalls_gt,
                          larges_table = larges_gt)
#Combine
data.tables %>% 
  gt() %>% 
  fmt_markdown(columns = TRUE) %>% 
  cols_label(smalls_table = "1-10 um",
             larges_table = "10-5,000 um") %>% 
  tab_source_note(md("Calculated from SSDs")) %>% 
  tab_header(title = "Microplastics HC5 (ug/L) by Data Collapse and Tier",
             subtitle = "Smalls and Larges") 
```

