---
title: "Stepwise Multiple Regressions"
author: "Leah Thornton Hampton"
date: "2/1/2021"
output:
  html_document: 
    code_folding: hide
    theme: sandstone
    toc: yes
    toc_float: yes
    toc_depth: 4
    number_sections: true
  word_document:
    toc: yes
---
```{r global-options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE)
```
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
#Packages
library(tidyverse)
library(MASS, exclude = 'select') #Stepwise regression package 
library(ggiraphExtra) #For plotting
library(ggeffects)
library(ggdark)
library(ggsci)
library(viridis)

require(readr)
#load aoc_z into dataframe. This file is generated from RDA_Maker.R
#source("Tox Data/RDA_Maker.R")
aoc_setup <- readRDS(file = "Tox Data/aoc_z.Rda")
```

This script runs multiple linear regressions and performs stepwise feature elimination.

# Set Up

```{r}

#Custom function that finds factors with only 1 level
#Source: https://stackoverflow.com/questions/44200195/how-to-debug-contrasts-can-be-applied-only-to-factors-with-2-or-more-levels-er

debug_contr_error <- function (dat, subset_vec = NULL) {
  if (!is.null(subset_vec)) {
    ## step 0
    if (mode(subset_vec) == "logical") {
      if (length(subset_vec) != nrow(dat)) {
        stop("'logical' `subset_vec` provided but length does not match `nrow(dat)`")
        }
      subset_log_vec <- subset_vec
      } else if (mode(subset_vec) == "numeric") {
      ## check range
      ran <- range(subset_vec)
      if (ran[1] < 1 || ran[2] > nrow(dat)) {
        stop("'numeric' `subset_vec` provided but values are out of bound")
        } else {
        subset_log_vec <- logical(nrow(dat))
        subset_log_vec[as.integer(subset_vec)] <- TRUE
        } 
      } else {
      stop("`subset_vec` must be either 'logical' or 'numeric'")
      }
    dat <- base::subset(dat, subset = subset_log_vec)
    } else {
    ## step 1
    dat <- stats::na.omit(dat)
    }
  if (nrow(dat) == 0L) warning("no complete cases")
  ## step 2
  var_mode <- sapply(dat, mode)
  if (any(var_mode %in% c("complex", "raw"))) stop("complex or raw not allowed!")
  var_class <- sapply(dat, class)
  if (any(var_mode[var_class == "AsIs"] %in% c("logical", "character"))) {
    stop("matrix variables with 'AsIs' class must be 'numeric'")
    }
  ind1 <- which(var_mode %in% c("logical", "character"))
  dat[ind1] <- lapply(dat[ind1], as.factor)
  ## step 3
  fctr <- which(sapply(dat, is.factor))
  if (length(fctr) == 0L) warning("no factor variables to summary")
  ind2 <- if (length(ind1) > 0L) fctr[-ind1] else fctr
  dat[ind2] <- lapply(dat[ind2], base::droplevels.factor)
  ## step 4
  lev <- lapply(dat[fctr], base::levels.default)
  nl <- lengths(lev)
  ## return
  list(nlevels = nl, levels = lev)
  }


```


#Variables to put into model

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Size (binned)
* Polymer type
* Shape
* Organism Group (this is being used as a temporary substitute for body size)
* Life Stage
* Level of Biological Organization
* Exposure Duration
* Acute/Chronic (this values are only complete for fish, molluscs, crustaceans and algae)
* Charge (negative/positive - categorical)
* Zeta potential 
* Particle Source (commercial, generated in-house, or mmodified commercial particles e.g., milling)

Response Variables
* Effect (Y/N) - Binary Data
* Effect (Y/N) - Binary Data x Effect Score (binned organism level effects)

#Continous Variable Distributions

There are several continuous variables that we want to feed into our model. Before doing so, we're going to check the distribution to see if any of them are skewed and need to be transformed. 

```{r, echo = FALSE}

#Dose - Count

Dose_Particles_mL <- aoc_setup$dose.particles.mL.master

hist(Dose_Particles_mL)

hist(log10(Dose_Particles_mL))

#Dose - Mass

Dose_Mass_L <- aoc_setup$dose.mg.L.master

hist(Dose_Mass_L)

hist(log10(Dose_Mass_L))

#Dose - Volume

`Dose_um^3_mL` <- aoc_setup$dose.um3.mL.master

hist(`Dose_um^3_mL`)

hist(log10(`Dose_um^3_mL`))

#Size - Continuous

Size <- aoc_setup$size.length.um.used.for.conversions

hist(Size)

hist(log10(Size))

#Exposure Duration

Exposure_Duration <- aoc_setup$exposure.duration.d

hist(Exposure_Duration)

hist(log10(Exposure_Duration))

#Zeta Potential

Zeta_Potential_mV <- aoc_setup$zetapotential.mV

hist(Zeta_Potential_mV)

#No need to log transform

aoc_setup_select <- aoc_setup %>% 
  mutate(log.dose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
  mutate(log.dose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(log.dose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(log.size.length.um.used.for.conversions = log10(size.length.um.used.for.conversions)) %>%
  mutate(log.exposure.duration.d = log10(exposure.duration.d)) %>% 
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    ))# %>%
  #mutate(effect_10 = factor(effect_10))


```

Due to skewed data, the following categories are log10 transformed before modeling: 

* Dose (counts)
* Dose (mass)
* Dose (volume)
* Size (continuous)
* Exposure Duration

# Model: Kitchen Sink

All Independent Variables, Response Variable: Effect (Y/N)

Full Model 

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
aoc_setup_select_1 <- aoc_setup_select  %>%
  filter(life_f != "Not Reported") %>% 
  filter(acute.chronic_f != "Unavailable") %>% 
  filter(size_f != "Not Reported") %>% 
  filter(poly_f != "Not Reported") %>%
  filter(shape_f != "Not Reported") %>%
  filter(environment != "Terrestrial") %>%
  dplyr::select(
    log.dose.particles.mL.master, 
    log.dose.mg.L.master,
    log.dose.um3.mL.master,
    log.size.length.um.used.for.conversions,
    # size_f,
    # poly_f,
    shape_f,
    org_f,
    life_f,
    bio_f,
    log.exposure.duration.d,
    acute.chronic_f,
    # charge_f,
    # zetapotential.mV,
    # particle_source_f,
    effect_10) 
#Check for single level factors
debug_contr_error(aoc_setup_select_1)

#Omit missing data before modeling
aoc_setup_select_1 <- na.omit(aoc_setup_select_1)

#Fit the full model
Full_Model <- lm(effect_10 ~., data = aoc_setup_select_1)

summary(Full_Model)
```

Stepwise Model - Both Directions

```{r, echo = FALSE}

#Stepwise model
step.model <- stepAIC(Full_Model, direction = "both", 
                      trace = FALSE)


summary(step.model)

```

# Model: Crustacean Fitness, Organism Level Endpoints Only
## Dose (mass)

Full Model 

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.mg.L.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)
```

```{r}
#Plots

ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = FALSE)
```

## Dose (volume)

Full Model 

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.um3.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)
```

```{r}
#Plots
ggPredict(m1_crust_model_dose,
          colorn=200,
 #         jitter=TRUE, 
point = FALSE,
          interactive = FALSE, 
show.summary = TRUE,
#digits = 3, 
se = FALSE)
```

```{r}
#particle couunt
#Select the columns that we want to feed into the model for simplicity
m1_crust_part <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.particles.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_particles <- glm(effect_10 ~ logdose.particles.mL.master * size.length.um.used.for.conversions, data = m1_crust_part, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_particles)
```
plot
```{r}
ggPredict(m1_crust_model_particles, colorn=1000, interactive = FALSE, jitter = FALSE)
```
### volume
```{r}
#particle couunt
#Select the columns that we want to feed into the model for simplicity
m1_crust_volume <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_volume <- glm(effect_10 ~ logdose.um3.mL.master * size.length.um.used.for.conversions, data = m1_crust_volume, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_volume)
```
plot
```{r}
ggPredict(m1_crust_model_volume, colorn=800, interactive = FALSE,jitter = FALSE)
```


```{r}
#alternative method using ggplot

#get probabilities
m1 <- m1_crust %>% 
  mutate(prob = predict(m1_crust_model_dose,
                        type = "response"))
  
# plot
ggplot(m1, aes(x = logdose.mg.L.master, y = effect_10, color = size.length.um.used.for.conversions))+
  geom_jitter() +
  geom_smooth(method = 'glm',
              aes(x = logdose.mg.L.master, y = prob), #use predicted NOEC probability
              method.args = list(family = binomial(link = 'logit')),
              se = FALSE, color = 'red') #+
  # geom_smooth(method = 'glm',
  #             method.args = list(family = binomial(link = 'probit')),
  #             se = FALSE, color = 'green') +
  # stat_smooth(method = "glm", method.args = list(family = "binomial"), se = FALSE)
```

```{r}
ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
```

## Dose (count)

Full Model 

```{r, echo = FALSE}

#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, logdose.particles.mL.master, logdose.um3.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.particles.mL.master*size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)

ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = FALSE)

```
## Dose (Surface Area)

Full Model 

```{r, echo = FALSE}
#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(logdose.surface.area.um2.mL.master = log10(dose.surface.area.um2.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.surface.area.um2.mL.master, size.length.um.used.for.conversions)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose_SA <- glm(effect_10 ~ logdose.surface.area.um2.mL.master * size.length.um.used.for.conversions, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose_SA)

ggPredict(m1_crust_model_dose_SA,colorn=700,jitter=FALSE, interactive = FALSE)
```

## Combo Dose Predictors

Full Model 

```{r, echo = FALSE}
#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(logdose.surface.area.um2.mL.master = log10(dose.surface.area.um2.mL.master)) %>%
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.surface.area.um2.mL.master, logdose.um3.mL.master)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_dose <- glm(effect_10 ~ logdose.surface.area.um2.mL.master , data = m1_crust,
                           na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_dose)

ggPredict(m1_crust_model_dose,colorn=700,jitter=FALSE, interactive = FALSE)
```


```{r}
require(ggiraphExtra)
#Plots
ggPredict(m1_crust_model_dose,colorn=100,jitter=FALSE, interactive = TRUE)
```

```{r eval=FALSE, include=FALSE}
#alternative method using ggpredict
require(ggeffects)
require(ggdark)
require(ggsci)
ggpredict(m1_crust_model_dose)
df <- ggpredict(m1_crust_model_dose, se = TRUE)
df <- ggpredict(m1_crust_model_dose_SA,terms=c("logdose.surface.area.um2.mL.master", "size.length.um.used.for.conversions"))

#note that faceting works best with discrete variable as interactive term
ggplot(df, aes(x = x, y = predicted, colour = group)) +
  geom_line() +
  geom_ribbon( aes(ymin = conf.low, ymax = conf.high, fill = group, color = NULL), alpha = .15) +
  facet_wrap(~.) +
  dark_theme_bw() +
  scale_color_tron()
```

```{r eval=FALSE, include=FALSE}
#alt method
plot(df, facet = TRUE)
```


## Polymers

```{r, echo = FALSE}
#Select the columns that we want to feed into the model for simplicity
poly <- aoc_z %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>%
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>%
  mutate(logdose.um3.mL.master = log10(dose.um3.mL.master)) %>%
  mutate(logdose.surface.area.um2.mL.master = log10(dose.surface.area.um2.mL.master)) %>%
  mutate(effect_10 = case_when( #convert ordinal to numeric
      effect == "Y" ~ 1,
      effect == "N" ~ 0
    ))
 # filter(acute.chronic_f == "Acute") %>% 
  #dplyr::select(c(effect_10, logdose.surface.area.um2.mL.master, max.size.ingest.mm, logdose.um3.mL.master)) %>% 
  #drop_na()

#Fit the full model
poly_glm <- glm(effect_10 ~ logdose.surface.area.um2.mL.master + poly_f, data = poly, na.action = "na.exclude", family = "binomial")

summary(poly_glm)

ggPredict(poly_glm,colorn=700,jitter=FALSE, interactive = FALSE)
```

## Discrete predictor and dose
```{r, echo = FALSE}
#Select the columns that we want to feed into the model for simplicity
m1_crust <- aoc_setup_select %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, size_f)) %>% 
  drop_na()

#Fit the full model
m1_crust_model_discrete <- glm(effect_10 ~ logdose.mg.L.master * size_f, data = m1_crust, na.action = "na.exclude", family = "binomial")

summary(m1_crust_model_discrete)
```
Plotted below.
```{r}
discrete <- ggpredict(m1_crust_model_discrete,terms = c("logdose.mg.L.master", "size_f"))
plot(discrete, facet = TRUE, add.data = TRUE,# colors = "quadro",
     alpha = 0.2, dodge = 0.2)
     #title = "All Aquatic Organisms")
```
#### Acute only crustacea
```{r, echo = FALSE}
require(arm) #bayesglm()
#Select the columns that we want to feed into the model for simplicity
m1_crust_acute <- aoc_setup %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(bio.org == "organism") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  filter(!effect_f == "NA") %>% 
  mutate(logdose.particles.mL.master = log10(dose.particles.mL.master)) %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, 
                  logdose.mg.L.master, 
                  logdose.particles.mL.master, 
                  size_f)) %>% 
  drop_na()

#Fit the full model
m1_crust_acute_model_discrete <- glm(effect_10 ~ 
                                      logdose.mg.L.master * size_f  + 
                                       logdose.particles.mL.master * size_f, 
                                     data = m1_crust_acute, 
                                     na.action = "na.exclude", 
                                     family = "binomial"#,
                                     #glm.control(maxit = 50)
                                     )

summary(m1_crust_acute_model_discrete)
```
The above glm is visualized below.
```{r}
acute_discrete <- ggpredict(m1_crust_acute_model_discrete,terms = c(#"logdose.mg.L.master", 
                                                                    "logdose.particles.mL.master",
                                                                    "size_f")) #ensure discrete factor is second
plot(acute_discrete)#, facet = TRUE, add.data = TRUE)
```

#### Alt Method using GG Plot
This approach achieves a similiar product as using the ggPredict() function, except it relies on ggplot(), which is more malleable and transparent. The general steps are to first create a new dataframe over 1000 values of size using expand.grid() then use predict() and plot() with geom_line() and colour=size.
```{r}
require(viridis)
#filtered dataset
#m1_crust
#model
#summary(m1_crust_model_dose )
#generate distribution of data
mockData <- expand.grid(size.length.um.used.for.conversions = seq(0.034, 5000, 1),
            logdose.particles.mL.master = seq(-4.195, 12.650,0.1))

mockData$effect_10 <- predict(m1_crust_model_dose,
                     mockData,
                     type = "response")
#plot
mockData %>% 
  ggplot(aes(x = logdose.particles.mL.master, 
             y = effect_10, 
             color = size.length.um.used.for.conversions)) +
  geom_line() +
  scale_color_gradientn(colors = topo.colors(7)) +
  #scale_color_viridis_c(option = "A") +
  dark_theme_bw()
```


## Survival package
### Cumulative Hazard by Time
```{r}
#include time
m1_crust <- aoc_setup %>% 
  filter(!effect_f == "NA") %>% 
  #filter(effect.metric == c("HONEC", "LOEC", "NOEC")) %>% 
  #filter(organism.group == "Crustacea") %>% 
  filter(!acute.chronic_f == "Unavailable") %>% 
  filter(bio.org == "organism") %>% 
  filter(!environment == "Terrestrial") %>% 
  filter(lvl1_f == "Fitness") %>% 
  filter(!size_f == "Not Reported") %>% 
  mutate(logdose.mg.L.master = log10(dose.mg.L.master)) %>% 
 # filter(acute.chronic_f == "Acute") %>% 
  dplyr::select(c(effect_10, logdose.mg.L.master, size_f, exposure.duration.d)) %>% 
  drop_na()

require(survival)
survival <- survival::coxph(survival::Surv(exposure.duration.d, effect_10) ~ logdose.mg.L.master + size_f, data = m1_crust)
#cumulative hazard
cumHaz <- ggpredict(survival,terms = c("logdose.mg.L.master", "size_f"), type = "cumhaz")
plot(cumHaz, facet = TRUE, add.data = TRUE, colors = "flat")
```
### Probability of Survival by Time

```{r}
#survival probability over time
pr <- ggpredict(survival, c("logdose.mg.L.master", "size_f"), type = "surv")
plot(pr, colors = "social")
```
