---
title: "Assessment Factors"
author: "Scott Coffin"
date: "3/30/2021"
output: 
  html_document:
    code_folding: hide
    theme: journal
    toc: yes
    toc_float: yes
    toc_depth: 6
    number_sections: true
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      warning=FALSE, message=FALSE,time_it = TRUE) #report
library(tidyverse)
library(ggpubr)
```

```{r include=FALSE}
require(readr)
#load aoc_z into dataframe. This file is generated from RDA_Maker.R
source("Tox Data/RDA_Maker.R")
#aoc_z <- readRDS("Tox Data/aoc_z.Rda") 
#must have "EValuate Chunks in directory: Project" in Global Options _> R Markdown in order for above line to work
```

# NOEC & LOEC Distribution

There doesn't seem to be much separation between acute and chronic NOEC/LOEC distributions. If this is the case, is an assessment factor even necessary?

Attempts to be more selective/curate the data further don't really seem to change anything (e.g., filtering for certain size fractions, using quality filters, etc.)

```{r NOEC & LOEC Distributions}

dist <- aoc_z # %>% 
 
  ##leachates already removed in RDAmaker.RmD
   #Remove leachate and additive/chemical transfer experiments
 # replace_na(list(chem.exp.typ.nominal = "Particle Only")) %>% 
#  dplyr::filter(leachate.only != "Y") %>%
 # mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
                                                #   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
                                                #   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
#  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") 
  # filter(tier_zero_tech_f == "Red Criteria Passed")  
  # filter(tier_zero_risk_f == "Red Criteria Passed")

#Dose in mass per volume 
Dist_NOEC_Mass <- dist %>% 
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Select NOECs
     filter(effect.metric == "NOEC") %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.mg.L.master) %>%  
     select(org_f, effect.metric, acute.chronic_f, dose.mg.L.master) %>%  
     ggplot(aes(x = dose.mg.L.master, fill = acute.chronic_f)) +
     geom_histogram(bins = 50)+
     facet_wrap(~org_f)+
     scale_x_log10() +
     labs(title = "Distribution of Derived NOEC Values", x = "Dose (mg/L)", y = "Count", fill = "Acute vs. Chronic")

plot(Dist_NOEC_Mass)

Dist_LOEC_Mass <- dist %>% 
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Select NOECs
     filter(effect.metric == "LOEC") %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.mg.L.master) %>% 
     select(org_f, effect.metric, acute.chronic_f, dose.mg.L.master) %>%  
     ggplot(aes(x = dose.mg.L.master, fill = acute.chronic_f)) +
     geom_histogram(bins = 50)+
     facet_wrap(~org_f)+
     scale_x_log10()+
     labs(title = "Distribution of Derived LOEC Values", x = "Dose (mg/L)", y = "Count", fill = "Acute vs. Chronic")

plot(Dist_LOEC_Mass)

#Dose in count per volume
Dist_NOEC_Count <- dist %>% 
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Select NOECs
     filter(effect.metric == "NOEC") %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.particles.mL.master) %>% 
     select(org_f, effect.metric, acute.chronic_f, dose.particles.mL.master) %>%  
     ggplot(aes(x = dose.particles.mL.master, fill = acute.chronic_f)) +
     geom_histogram(bins = 50)+
     facet_wrap(~org_f)+
     scale_x_log10() +
     labs(title = "Distribution of Derived NOEC Values", x = "Dose (particles/mL)", y = "Count", fill = "Acute vs. Chronic")

plot(Dist_NOEC_Count)

Dist_LOEC_Count <- dist %>% 
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Select NOECs
     filter(effect.metric == "LOEC") %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.particles.mL.master) %>% 
     select(org_f, effect.metric, acute.chronic_f, dose.particles.mL.master) %>%  
     ggplot(aes(x = dose.particles.mL.master, fill = acute.chronic_f)) +
     geom_histogram(bins = 50)+
     facet_wrap(~org_f)+
     scale_x_log10()+
     labs(title = "Distribution of Derived LOEC Values", x = "Dose (particles/mL)", y = "Count", fill = "Acute vs. Chronic")

plot(Dist_LOEC_Count)

```

```{r How many species within each taxa have both acute and chronic values? }

Species_Taxa_Mass <- aoc_z %>% 
     #Select NOECs
     filter(effect.metric == c("NOEC", "LOEC")) %>%
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.mg.L.master) %>%
     #Group the data by organism group and acute vs chronic data 
     group_by(org_f, species_f, acute.chronic_f, effect.metric) %>%
     #Create a new column for the average NOEC
     summarise() %>% 
     ungroup() %>% 
     group_by(org_f, acute.chronic_f, effect.metric) %>% 
     mutate(species_count = n_distinct(species_f)) %>% 
     ungroup() %>% 
     distinct(org_f, acute.chronic_f, effect.metric, species_count)

Species_Taxa_Mass
```

```{r NOECs}

#Compare differences between acute and chronic values for NOECs in each organism group

Mean_NOEC <- aoc_z %>% 
     #Select NOECs
     filter(effect.metric == "NOEC") %>%
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.mg.L.master) %>%
     #Group the data by organism group and acute vs chronic data 
     group_by(org_f, acute.chronic_f) %>%
     #Create a new column for the average NOEC
     summarise(averageNOEC = mean(dose.mg.L.master)) 
     
acute <- Mean_NOEC %>% 
     #Select only acute data
     filter(acute.chronic_f == "Acute") %>% 
     rename(averageNOEC_acute = averageNOEC)
     
chronic <- Mean_NOEC %>% 
     #Select only chronic data
     filter(acute.chronic_f == "Chronic") %>% 
     rename(averageNOEC_chronic = averageNOEC)

acute_chronic <- left_join(acute, chronic, by = c("org_f" = "org_f")) %>% 
  mutate(change_NOEC = averageNOEC_acute/averageNOEC_chronic)

#Regression between acute and chronic NOECs

plot <- acute_chronic %>% 
  ggplot(aes(x = log10(averageNOEC_chronic), y = log10(averageNOEC_acute))) +
  geom_point() 

plot(plot)
```

```{r LOECs}

#Compare differences between acute and chronic values for NOECs in each organism group

Mean_LOEC <- aoc_z %>% 
     #Select NOECs
     filter(effect.metric == "LOEC") %>%
     #Select data where acute and chronic exposure is defined
     filter(acute.chronic_f == c("Acute", "Chronic")) %>%
     #Drop rows where there is no dosing data in mass per volume
     drop_na(dose.mg.L.master) %>%
     #Group the data by organism group and acute vs chronic data 
     group_by(org_f, acute.chronic_f) %>%
     #Create a new column for the average LOEC
     summarise(averageLOEC = mean(dose.mg.L.master)) 
     
acute <- Mean_LOEC %>% 
     #Select only acute data
     filter(acute.chronic_f == "Acute") %>% 
     rename(averageLOEC_acute = averageLOEC)
     
chronic <- Mean_LOEC %>% 
     #Select only chronic data
     filter(acute.chronic_f == "Chronic") %>% 
     rename(averageLOEC_chronic = averageLOEC)

acute_chronic <- left_join(acute, chronic, by = c("org_f" = "org_f")) %>% 
  mutate(change_LOEC = averageLOEC_acute/averageLOEC_chronic)

#Regression between acute and chronic LOECs

plot <- acute_chronic %>% 
  ggplot(aes(x = log10(averageLOEC_chronic), y = log10(averageLOEC_acute))) +
  geom_point() 

plot(plot)
```


# Data Filtering and Grouping

```{r}
# collapse data (average) for all combinations of selected variables
grouped <- aoc_z %>% 
   filter(tier_zero_tech_f == "Red Criteria Passed") %>% 
   filter(tier_zero_risk_f == "Red Criteria Passed") %>% 
  #Select NOECs/LOECs
  #filter(effect.metric == c("NOEC", "LOEC")) %>%
  #Select data where acute and chronic exposure is defined
  filter(acute.chronic_f == c("Acute", "Chronic")) %>%
  filter(size_f != "Not Reported") %>% 
  #Drop rows where there is no dosing data in mass per volume
 # drop_na(dose.mg.L.master) %>%
  #Group the data by organism group and acute vs chronic data 
  group_by(org_f, acute.chronic_f, effect.metric, lvl1_f, size_f) %>% 
           #size_f) 
  droplevels() %>% 
  #Create a new column for the average effect for each combination of the matrix
  summarise(meanEffect.particles = mean(dose.particles.mL.master),
            meanEffect.mass = mean(dose.mg.L.master),
            meanEffect.volume = mean(dose.um3.mL.master),
            sdEffect.mass = sd(dose.mg.L.master),
            nstudies = n()) #%>% 
  #drop_na()

grouped
```
## NOEC/LOEC
```{r}
#create unique ID for each combo of interest for NOECs and LOECs
NOEC <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f, 
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric == "NOEC") %>% 
  mutate(effectNOEC.particles = meanEffect.particles,
         effectNOEC.mass = meanEffect.mass,
         effectNOEC.volume = meanEffect.volume,
         sdNOEC.mass = sdEffect.mass) %>% 
  mutate(stderrorNOEC.mass = sdNOEC.mass / nstudies)

LOEC <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f,
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric == "LOEC") %>% 
    mutate(effectLOEC.particles = meanEffect.particles,
         effectLOEC.mass = meanEffect.mass,
         effectLOEC.volume = meanEffect.volume,
         sdLOEC.mass = sdEffect.mass) %>% 
  mutate(stderrorLOEC.mass = sdLOEC.mass / nstudies)

# join NOEC and LOEC dataframes and compute LOEC/NOEC ratio for each combo
NOEC.LOEC <- full_join(NOEC, LOEC, by = "combo") %>%
  select(c(combo,effectNOEC.particles, 
           effectLOEC.particles, 
           effectNOEC.mass, 
           effectLOEC.mass, 
           effectNOEC.volume,
           effectLOEC.volume,
           size_f.x,
           size_f.y,
           nstudies.x, 
           nstudies.y,
           sdLOEC.mass,
           sdNOEC.mass,
           stderrorNOEC.mass,
           stderrorLOEC.mass)) %>% 
  mutate(LOEC.NOEC.ratio.mass = effectLOEC.mass / effectNOEC.mass) %>%
  mutate(LOEC.NOEC.ratio.particles = effectLOEC.particles / effectNOEC.particles) %>% 
  mutate(LOEC.NOEC.ratio.volume = effectLOEC.volume / effectNOEC.volume) %>%
  filter(nstudies.y > 1) %>% 
  filter(nstudies.x > 1) %>% 
  #filter(LOEC.NOEC.ratio.mass > 1) %>%
  #filter(LOEC.NOEC.ratio.particles > 1) %>% 
  drop_na()

NOEC.LOEC
```
### Mass
```{r}
noec.loec.lm <- summary(lm(effectLOEC.mass ~ effectNOEC.mass, data = NOEC.LOEC))
noec.loec.lm
```
### Particle Count

```{r}
noec.loec.lm.particle <- summary(lm(effectLOEC.particles ~ effectNOEC.particles, data = NOEC.LOEC))
noec.loec.lm.particle
```

### Volume
```{r}
noec.loec.lm.volume <- summary(lm(effectLOEC.volume ~ effectNOEC.volume, data = NOEC.LOEC[-3,]))
noec.loec.lm.volume
```

```{r}
NOEC.LOEC %>% 
  ggplot(aes(x = effectNOEC.mass, y = effectLOEC.mass)) +
  geom_point() +
  geom_errorbar(aes(ymin = effectLOEC.mass - stderrorLOEC.mass, ymax = effectLOEC.mass + stderrorLOEC.mass,
                    xmin = effectNOEC.mass - stderrorNOEC.mass, xmax = effectNOEC.mass + stderrorNOEC.mass)) +
  geom_smooth(method = "lm") +
  stat_regline_equation(label.x = 1, label.y = 35) +
  stat_cor(label.x = 1, label.y = 30) +
  geom_text(x = 5, y= 40, color = "forestgreen",
    label = paste("NOEC/LOEC AF:",
    round(noec.loec.lm$coefficients[2,1],2), #coefficient
            "+-",
            round(noec.loec.lm$coefficients[2,2],2), #std error
    "(mean +- std.error)"))+
  ggrepel::geom_text_repel(aes(label = combo), color = "red") +
  labs(title = "LOEC/NOEC Assessment Factor Derivation",
       subtitle = "Grouped (averaged) by taxa, acute/chronic, particle size bin, level 1 effect",
       caption = "Data filtered for red criteria. Aquatic Tox database. 03-30-2021.") +
  xlab("Average NOEC (mg/L)") +
  ylab("Average LOEC (mg/L)") +
  theme_minimal() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

## HONEC/NOEC
```{r}
#create unique ID for each combo of interest for NOECs and HONECs
NOEC <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f, 
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric == "NOEC") %>% 
  mutate(effectNOEC.particles = meanEffect.particles,
         effectNOEC.mass = meanEffect.mass,
         effectNOEC.volume = meanEffect.volume,
         sdNOEC.mass = sdEffect.mass) %>% 
  mutate(stderrorNOEC.mass = sdNOEC.mass / nstudies)

HONEC <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f,
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric == "HONEC") %>% 
    mutate(effectHONEC.particles = meanEffect.particles,
         effectHONEC.mass = meanEffect.mass,
         effectHONEC.volume = meanEffect.volume,
         sdHONEC.mass = sdEffect.mass) %>% 
  mutate(stderrorHONEC.mass = sdHONEC.mass / nstudies)

# join NOEC and HONEC dataframes and compute HONEC/NOEC ratio for each combo
NOEC.HONEC <- full_join(NOEC, HONEC, by = "combo") %>%
  select(c(combo,effectNOEC.particles, 
           effectHONEC.particles, 
           effectNOEC.mass, 
           effectHONEC.mass, 
           effectNOEC.volume, 
           effectHONEC.volume, 
           size_f.x,
           size_f.y,
           nstudies.x, 
           nstudies.y,
           sdHONEC.mass,
           sdNOEC.mass,
           stderrorNOEC.mass,
           stderrorHONEC.mass)) %>% 
  mutate(HONEC.NOEC.ratio.mass = effectHONEC.mass / effectNOEC.mass) %>%
  mutate(HONEC.NOEC.ratio.particles = effectHONEC.particles / effectNOEC.particles) %>% 
  mutate(HONEC.NOEC.ratio.volume = effectHONEC.volume / effectNOEC.volume) %>% 
  filter(nstudies.y > 1) %>% 
  filter(nstudies.x > 1) %>% 
  #filter(HONEC.NOEC.ratio.mass > 1) %>%
  #filter(HONEC.NOEC.ratio.particles > 1) %>% 
  drop_na()

NOEC.HONEC
```

### Mass
```{r}
honec.loec.lm <- summary(lm(effectHONEC.mass ~ effectNOEC.mass, data = NOEC.HONEC))
honec.loec.lm
```

### Particle Count

```{r}
honec.noec.lm <- summary(lm(effectHONEC.particles ~ effectNOEC.particles, data = NOEC.HONEC))
honec.noec.lm
```
### Volume
```{r}
honec.noec.lm.volume <- summary(lm(effectHONEC.volume ~ effectNOEC.volume, data = NOEC.HONEC[-3:-4,]))
honec.noec.lm.volume
```

```{r}
NOEC.HONEC %>% 
  ggplot(aes(x = effectNOEC.mass, y = effectHONEC.mass)) +
  geom_point() +
  geom_errorbar(aes(ymin = effectHONEC.mass - stderrorHONEC.mass, ymax = effectHONEC.mass + stderrorHONEC.mass,
                    xmin = effectNOEC.mass - stderrorNOEC.mass, xmax = effectNOEC.mass + stderrorNOEC.mass)) +
  geom_smooth(method = "lm") +
  stat_regline_equation(label.x = 1, label.y = 150) +
  stat_cor(label.x = 1, label.y = 135) +
  geom_text(x = 25, y= 155,color = "forestgreen",
    label = paste("HONEC/NOEC AF:",
    round(honec.noec.lm$coefficients[2,1],2), #coefficient
            "+-",
            round(honec.noec.lm$coefficients[2,2],2), #std error
    "(mean +- std.error)"))+
  ggrepel::geom_text_repel(aes(label = combo), color = "red") +
  labs(title = "HONEC/NOEC Assessment Factor Derivation",
       subtitle = "Grouped (averaged) by taxa, acute/chronic, particle size bin, level 1 effect",
       caption = "Data filtered for red criteria. Aquatic Tox database. 03-30-2021.") +
  xlab("Average NOEC (mg/L)") +
  ylab("Average HONEC (mg/L)") +
  theme_minimal() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

## EC50/NOEC
```{r}
#create unique ID for each combo of interest for NOECs and HONECs
NOEC <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f, 
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric == "NOEC") %>% 
  mutate(effectNOEC.particles = meanEffect.particles,
         effectNOEC.mass = meanEffect.mass,
         effectNOEC.volume = meanEffect.volume,
         sdNOEC.mass = sdEffect.mass) %>% 
  mutate(stderrorNOEC.mass = sdNOEC.mass / nstudies)

EC50 <- grouped %>% 
  mutate(combo = paste(org_f, acute.chronic_f,
                       size_f,
                       lvl1_f)) %>% 
  filter(effect.metric %in% c("EC50",
                              "EC10", 
                              "EMT50" ,
                              "LC50"
                              )) %>% 
    mutate(effectEC50.particles = meanEffect.particles,
         effectEC50.mass = meanEffect.mass,
         effectEC50.volume = meanEffect.volume,
         sdEC50.mass = sdEffect.mass) %>% 
  mutate(stderrorEC50.mass = sdEC50.mass / nstudies)

# join NOEC and EC50 dataframes and compute EC50/NOEC ratio for each combo
NOEC.EC50 <- full_join(NOEC, EC50, by = "combo") %>%
  select(c(combo,effectNOEC.particles, 
           effectEC50.particles, 
           effectNOEC.mass, 
           effectEC50.mass, 
           effectNOEC.volume, 
           effectEC50.volume, 
           size_f.x,
           size_f.y,
           nstudies.x, 
           nstudies.y,
           sdEC50.mass,
           sdNOEC.mass,
           stderrorNOEC.mass,
           stderrorEC50.mass)) %>% 
  mutate(EC50.NOEC.ratio.mass = effectEC50.mass / effectNOEC.mass) %>%
  mutate(EC50.NOEC.ratio.particles = effectEC50.particles / effectNOEC.particles) %>% 
  mutate(EC50.NOEC.ratio.volume = effectEC50.volume / effectNOEC.volume) %>%
  filter(nstudies.y > 1) %>% 
  filter(nstudies.x > 1) %>% 
  #filter(EC50.NOEC.ratio.mass > 1) %>%
  #filter(EC50.NOEC.ratio.particles > 1) %>% 
  drop_na()

NOEC.EC50
```
### Mass
```{r}
EC50.noec.lm <- summary(lm(effectEC50.mass ~ effectNOEC.mass, data = NOEC.EC50))
EC50.noec.lm
```
### Volume
```{r}
EC50.noec.lm.volume <- summary(lm(effectEC50.volume ~ effectNOEC.volume, data = NOEC.EC50))
EC50.noec.lm.volume
```


```{r}
NOEC.EC50 %>% 
  ggplot(aes(x = effectNOEC.mass, y = effectEC50.mass)) +
  geom_point() +
  geom_errorbar(aes(ymin = effectEC50.mass - stderrorEC50.mass, ymax = effectEC50.mass + stderrorEC50.mass,
                    xmin = effectNOEC.mass - stderrorNOEC.mass, xmax = effectNOEC.mass + stderrorNOEC.mass)) +
  geom_smooth(method = "lm") +
  stat_regline_equation(label.x = 1, label.y = 150) +
  stat_cor(label.x = 1, label.y = 135) +
  geom_text(x = 25, y= 155,color = "forestgreen",
    label = paste("EC50/NOEC AF:",
    round(EC50.noec.lm$coefficients[2,1],2), #coefficient
            "+-",
            round(EC50.noec.lm$coefficients[2,2],2), #std error
    "(mean +- std.error)"))+
  ggrepel::geom_text_repel(aes(label = combo), color = "red") +
  labs(title = "EC50/NOEC Assessment Factor Derivation",
       subtitle = "Grouped (averaged) by taxa, acute/chronic, particle size bin, level 1 effect",
       caption = "Data filtered for red criteria. Aquatic Tox database. 03-30-2021.") +
  xlab("Average NOEC (mg/L)") +
  ylab("Average EC50 (mg/L)") +
  theme_minimal() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```

# Acute/Chronic
## Grouping
```{r}
# collapse data (average) for all combinations of selected variables
grouped2 <- aoc_z %>% 
   filter(tier_zero_tech_f == "Red Criteria Passed") %>% 
   filter(tier_zero_risk_f == "Red Criteria Passed") %>% 
  #Select NOECs/LOECs
  filter(effect.metric == c("NOEC", "LOEC")) %>%
  #Select data where acute and chronic exposure is defined
  filter(acute.chronic_f == c("Acute", "Chronic")) %>%
  filter(size_f != "Not Reported") %>% 
  #Drop rows where there is no dosing data in mass per volume
 # drop_na(dose.mg.L.master) %>%
  #Group the data by organism group and acute vs chronic data 
  group_by(org_f, 
           acute.chronic_f, 
           effect.metric,
           #size_f,
           lvl1_f) %>% 
           #size_f) 
  droplevels() %>% 
  #Create a new column for the average effect for each combination of the matrix
  summarise(meanEffect.particles = mean(dose.particles.mL.master),
            meanEffect.mass = mean(dose.mg.L.master),
            sdEffect.mass = sd(dose.mg.L.master),
            nstudies = n()) #%>% 
  #drop_na()

grouped2
```

## Subsetting
```{r}
#create unique ID for each combo of interest for NOECs and LOECs
Acute <- grouped2 %>% 
  mutate(combo = paste(org_f, 
                       effect.metric,
                       #size_f,
                       lvl1_f)) %>% 
  filter(acute.chronic_f == "Acute") %>% 
  mutate(effectAcute.particles = meanEffect.particles,
         effectAcute.mass = meanEffect.mass,
         sdAcute.mass = sdEffect.mass) %>% 
  mutate(stderrorAcute.mass = sdAcute.mass / nstudies)

Chronic <- grouped2 %>% 
  mutate(combo = paste(org_f, 
                       effect.metric,
                      # size_f,
                       lvl1_f)) %>% 
  filter(acute.chronic_f == "Chronic") %>% 
    mutate(effectChronic.particles = meanEffect.particles,
         effectChronic.mass = meanEffect.mass,
         sdChronic.mass = sdEffect.mass) %>% 
  mutate(stderrorChronic.mass = sdChronic.mass / nstudies)

# join Acute and Chronic dataframes and compute Chronic/Acute ratio for each combo
Acute.Chronic <- full_join(Acute, Chronic, by = "combo") %>%
  select(c(combo,effectAcute.particles, 
           effectChronic.particles, 
           effectAcute.mass, 
           effectChronic.mass, 
          # size_f.x,
           #size_f.y,
           nstudies.x, 
           nstudies.y,
           sdChronic.mass,
           sdAcute.mass,
           stderrorAcute.mass,
           stderrorChronic.mass)) %>% 
  mutate(Chronic.Acute.ratio.mass = effectChronic.mass / effectAcute.mass) %>%
  mutate(Chronic.Acute.ratio.particles = effectChronic.particles / effectAcute.particles) %>% 
  #filter(nstudies.y > 1) %>% 
  #filter(nstudies.x > 1) %>% 
  filter(Chronic.Acute.ratio.mass > 1) %>%
  filter(Chronic.Acute.ratio.mass < 1000) %>% 
  drop_na(effectAcute.mass, effectChronic.mass)

Acute.Chronic
```
```{r}
acute.chronic.lm <- summary(lm(effectChronic.mass ~ effectAcute.mass, data = Acute.Chronic))
acute.chronic.lm
```

```{r}
Acute.Chronic %>% 
  ggplot(aes(x = effectAcute.mass, y = effectChronic.mass)) +
  geom_point() +
  geom_errorbar(aes(ymin = effectChronic.mass - stderrorChronic.mass, ymax = effectChronic.mass + stderrorChronic.mass,
                    xmin = effectAcute.mass - stderrorAcute.mass, xmax = effectAcute.mass + stderrorAcute.mass)) +
  geom_smooth(method = "lm", se = FALSE) +
  stat_regline_equation(label.x = 5, label.y = 50) +
  stat_cor(label.x = 5, label.y = 30) +
   geom_text(x = 5, y= 150,color = "forestgreen",
    label = paste("Acute/Chronic AF:",
    round(acute.chronic.lm$coefficients[2,1],2), #coefficient
            "+-",
            round(acute.chronic.lm$coefficients[2,2],2), #std error
    "(mean +- std.error)"))+
  ggrepel::geom_text_repel(aes(label = combo), color = "red") +
  labs(title = "Chronic/Acute Assessment Factor Derivation",
       subtitle = "Grouped (averaged) by taxa, Acute/chronic, particle size bin, level 1 effect",
       caption = "Data filtered for red criteria. Aquatic Tox database. 03-30-2021.") +
  xlab("Average Acute (mg/L)") +
  ylab("Average Chronic (mg/L)") +
  theme_minimal() +
  theme(legend.position = "none",
    plot.title = element_text(hjust = 0.5, size = 14),
        axis.title = element_text(size = 16),
        axis.text =  element_text(size = 16),
        legend.text = element_text(size =14),
        legend.title = element_blank(),
        plot.subtitle = element_text(hjust = 0.5, size = 14))
```


## Multiple regression approach
```{r}
#alt approach 
filtered <- aoc_z %>% 
  #Remove leachate and additive/chemical transfer experiments
  replace_na(list(chem.exp.typ.nominal = "Particle Only")) %>% 
  dplyr::filter(leachate.only != "Y") %>%
  mutate(chem.exp.typ.nominal_f = factor(case_when(chem.exp.typ.nominal == "Particle Only" ~ "Particle Only",
                                                   chem.exp.typ.nominal == "co.exp" ~ "Chemical Co-Exposure",
                                                   chem.exp.typ.nominal == "sorbed" ~ "Chemical Transfer"))) %>% 
  dplyr::filter(chem.exp.typ.nominal_f == "Particle Only") %>% 
  #Select NOECs/LOECs
  filter(effect.metric == c("NOEC", "LOEC")) %>%
  #Select data where acute and chronic exposure is defined
  filter(acute.chronic_f == c("Acute", "Chronic")) %>%
  filter(size_f != "Not Reported") %>% 
  #Drop rows where there is no dosing data in mass per volume
  drop_na(dose.mg.L.master) %>%
  droplevels()

summary(lm(dose.mg.L.master ~ effect.metric + lvl2_f + size_f + species_f + life.stage, 
   data = filtered))
```
```{r}
filtered %>% 
  ggplot(aes(y = dose.mg.L.master, x = effect.metric)) +
  geom_boxplot() +
  scale_y_log10()
```

